---
title: "Restaurant EDA"
date: '`r Sys.Date()`'
output:
  html_document:
    number_sections: true
    fig_caption: true
    toc: true
    fig_width: 7
    fig_height: 4.5
    theme: cosmo
    highlight: tango
    code_folding: hide
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo=TRUE, error=FALSE)
```

# Introduction {.tabset .tabset-fade .tabset-pills}
This tutorial is a replicate of kaggle kernel writen by [Heads or Tails](https://www.kaggle.com/headsortails).
The goal of this EDA is to predict the statistics of visotors in two touristics areas in Japon. The challenge was published by kaggle.

## Load librairies

```{r}
#library(readr)
library(tibble)
library('stringr') # string manipulation
library('data.table') # data manipulation
library(dplyr)
library(lubridate)
library(ggplot2)
library(grid)
library(leaflet)
library(tidyr)
library(ggExtra)
library(ggridges)
```


## Load data

```{r}
air_visit <- as_tibble(fread(str_c("air_visit_data.csv")))
air_reserve <- as_tibble(fread(str_c('air_reserve.csv'))) 
air_store_info <- as_tibble(fread(str_c('air_store_info.csv')))
date_info <- as_tibble(fread(str_c('date_info.csv')))
hpg_reserve <- as_tibble(fread(str_c('hpg_reserve.csv')))
hpg_store_info <- as_tibble((fread(str_c('hpg_store_info.csv'))))
sample_submission <- as_tibble(fread(str_c('sample_submission.csv')))
store_id_relation <- as_tibble(fread(str_c('store_id_relation.csv')))
```

## multiplot function

```{r}
# Define multiple plot function
#
# ggplot objects can be passed in ..., or to plotlist (as a list of ggplot objects)
# - cols:   Number of columns in layout
# - layout: A matrix specifying the layout. If present, 'cols' is ignored.
#
# If the layout is something like matrix(c(1,2,3,3), nrow=2, byrow=TRUE),
# then plot 1 will go in the upper left, 2 will go in the upper right, and
# 3 will go all the way across the bottom.
#
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {

  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}
```

# Overview: File structure and content {.tabset .tabset-fade .tabset-pills}

## Air visits
```{r}
summary(air_visit)
```


```{r}
tibble::glimpse(air_visit)
```

The tibble shows the number of visitors by store for each date.
```{r}
air_visit %>% dplyr::distinct(air_store_id) %>% nrow()
air_visit %>% dplyr::distinct(visit_date) %>% nrow()
```

There is 829 different store and 478 dates

## Air Reserve

```{r}
summary(air_reserve)
```

```{r}
glimpse(air_reserve)
```
The number of reservation is 92378, distributed for 314 air store.

```{r}
air_reserve %>% dplyr::distinct(air_store_id) %>% nrow()
```

## HPG Reserve

```{r}
summary(hpg_reserve)
```

```{r}
tibble::glimpse(hpg_reserve)
```

There are 2000320 revervation for 13325 stores

```{r}
hpg_reserve %>% dplyr::distinct(hpg_store_id) %>% nrow()
```

## Air Store

```{r}
summary(air_store_info)
```

```{r}
tibble::glimpse(air_store_info)
```

The number of air area
```{r}
air_store_info %>% dplyr::distinct(air_area_name) %>% nrow()
```

The number of stores

```{r}
air_store_info %>% dplyr::distinct(air_store_id) %>% nrow()
```

The number of latitude

```{r}
air_store_info %>% dplyr::distinct(latitude) %>% nrow()
```
The number of longitude
```{r}
air_store_info %>% dplyr::distinct(longitude) %>% nrow()
```

the number of cuisine genre in air

```{r}
air_store_info %>% dplyr::distinct(air_genre_name) %>% nrow()
```

## HPG Stores

```{r}
summary(hpg_store_info)
```

```{r}
tibble::glimpse(hpg_store_info)
```
There are 4690 stores

```{r}
hpg_store_info %>% dplyr::distinct(latitude) %>% nrow()
```

There are 119 area name
```{r}
hpg_store_info %>% dplyr::distinct(hpg_area_name) %>% nrow()
```

There are  34 genres of cuisine in HPG

```{r}
hpg_store_info %>% dplyr::distinct(hpg_genre_name) %>% nrow()
```

## Holidayes

```{r}
glimpse(date_info)
```

## Store IDs

```{r}
glimpse(store_id_relation)
```
There are 150 relation between Air (829 stores) and HPG (4690 stores) stores.

## Sample Submission

```{r}
tibble::glimpse(sample_submission)
```

the submision file should be two columns. The first one is the cancatenation of store_id and date. the 2sd is the number of visitors.

## Missing Values


```{r}
purrr::map(ls(), is.na)
```

## Reformating dates
```{r}
air_visit <- air_visit %>% mutate(visit_date = ymd(visit_date))
air_reserve <- air_reserve %>% mutate(visit_datetime = ymd_hms(visit_datetime),
                                      reserve_datetime = ymd_hms(reserve_datetime))
hpg_reserve <- hpg_reserve %>% mutate(visit_datetime = ymd_hms(visit_datetime),
                                      reserve_datetime = ymd_hms(reserve_datetime))

air_store_info <- air_store_info %>% mutate(air_genre_name = as.factor(air_genre_name),
                                            air_area_name = as.factor(air_area_name))
hpg_store_info <- hpg_store_info %>% mutate(hpg_genre_name = as.factor(hpg_genre_name),
                                            hpg_area_name = as.factor(hpg_area_name))

date_info <- date_info %>% mutate(holiday_flg = as.logical(holiday_flg),
                                  date = ymd(calendar_date))
```

# Individual feature visualisations {.tabset .tabset-fade .tabset-pills}
Here we have a first look at the distribution of the feature in our individual data files before combining them for a more detailed analysis.

## Air Visits
```{r}
 p1 <- air_visit %>% 
  group_by(visit_date) %>% 
  summarise(all_visitors = sum(visitors)) %>%
  ggplot(aes(visit_date, all_visitors)) +
  geom_line(col = "blue") +
  labs(y = "All visitors", x = "Date")

 p2 <- air_visit %>%
   ggplot(aes(visitors)) +
   geom_vline(xintercept = 20, color ="red") +
   geom_histogram(fill = "blue", bins = 30) +
   #xlim(0,150)
   scale_x_log10()
 
 p3 <- air_visit %>%
   mutate(wday = wday(visit_date, label = TRUE)) %>%
   group_by(wday) %>%
   summarise(visits = median(visitors)) %>%
   ggplot(aes(wday, visits, fill = wday)) +
   geom_col() +
   theme(legend.position = "none") +
   labs(x = "Day of the week", y = "Median Visitors")
 
 p4  <- air_visit %>%
   mutate(Month = month(visit_date, label = TRUE)) %>%
   group_by(Month) %>%
   summarise(visits = median(visitors)) %>%
   ggplot(aes(Month, visits, fill = Month)) +
   geom_col() +
   theme(legend.position = 'none') +
   labs(x = "MOnth", y = "Median Visitors")
 
 layout <- matrix(c(1,1,1,1,2,3,4,4), 2, 4, byrow = TRUE)
 multiplot(p1, p2, p3, p4, layout=layout)
```


We will be forecasting for the last week of April plus May 2017, so let’s look at this time range in our 2016 training data:

```{r}
air_visit %>%
  filter(visit_date > ymd("2016-04-15") & visit_date < ymd("2016-06-15")) %>%
  group_by(visit_date) %>%
  summarise(all_visitors = sum(visitors)) %>%
  ggplot(aes(visit_date, all_visitors)) +
  geom_line() +
  geom_smooth(method = "loess", color = "blue", span = 1/10) +
  labs(y = "All visitors", x = "Date")
```

## Air Reservations

```{r}
foo <- air_reserve %>%
  mutate(reserve_date = date(reserve_datetime),
         reserve_hour = hour(reserve_datetime),
         reserve_wday = wday(reserve_datetime, label = TRUE),
         visit_date = date(visit_datetime),
         visit_hour = hour(visit_datetime),
         visit_wday = wday(visit_datetime, label = TRUE),
         diff_hour = time_length(visit_datetime - reserve_datetime, unit = "hour"),
         diff_day = time_length(visit_datetime - reserve_datetime, unit = "day")) 

  p1 <- foo %>%
  group_by(visit_date) %>%
  summarise(all_vitisors = sum(reserve_visitors)) %>%
  ggplot(aes(visit_date, all_vitisors)) +
  geom_line() +
  labs(x = "Air visit date")
  
  p2 <- foo %>%
    group_by(visit_hour) %>%
    summarise(all_visitors = sum(reserve_visitors)) %>%
    ggplot(aes(visit_hour, all_visitors)) +
    geom_col(fill = "blue") +
    labs(x = "Time from reservation to visit [hours]")
  
  p3 <- foo %>%
    filter(diff_hour < 24 *7) %>%
    group_by(diff_hour) %>%
    summarise(all_visitors = sum(reserve_visitors)) %>%
    ggplot(aes(diff_hour, all_visitors)) +
    geom_col(fill= "blue") +
    labs(x = "Time from reservation (hours)", y = "all_visitors")
  
  layout <- matrix(c(1, 1, 2, 3), 2, 2, byrow = TRUE)
  
  multiplot(p1, p2, p3, layout = layout)
```

## HPG Reservation

```{r}
too <- hpg_reserve %>%
  mutate(visit_date = date(visit_datetime),
         visit_hour = hour(visit_datetime),
         visit_wday = wday(visit_datetime),
         reserve_date = date(reserve_datetime),
         reserve_hour = hour(reserve_datetime),
         reserve_wday = wday(reserve_datetime),
         diff_hour = time_length(visit_datetime - reserve_datetime, unit = "hour"),
         diff_day = time_length(visit_datetime - reserve_datetime, unit = "day")
         )

p1 <- too %>%
  group_by(visit_date) %>%
  summarise(all_visitors = sum(reserve_visitors)) %>%
  ggplot(aes(visit_date, all_visitors)) +
  geom_line()

p2 <- too %>%
  group_by(visit_hour) %>%
  summarise(all_visitors = sum(reserve_visitors)) %>%
  ggplot(aes(visit_hour, all_visitors)) +
  geom_col(fill="red")

p3 <- too %>%
  group_by(diff_hour) %>%
  summarise(all_visitors = sum(reserve_visitors)) %>%
  ggplot(aes(diff_hour, all_visitors)) +
  geom_col(fill = "red") +
  xlim(0,150)

 layout <- matrix(c(1,1,2,3), 2, 2, byrow = TRUE)
 multiplot(p1, p2, p3, layout = layout)
```

## Air Store

```{r}
leaflet(air_store_info) %>%
  addTiles() %>%
  addProviderTiles("CartoDB.Positron") %>%
  addMarkers(~longitude, ~latitude,
             popup = ~air_store_id, label = ~air_genre_name,
             clusterOptions = markerClusterOptions()
             )
```


```{r}
p1 <- air_store_info %>%
  group_by(air_area_name) %>%
  count() %>%
  ungroup() %>%
  top_n(15,n) %>%
  ggplot(aes(reorder(air_area_name, n, FUN = min), n, fill = air_area_name )) +
  geom_col()+
  coord_flip()+
  theme(legend.position = "none") +
  labs(x= "Top 15 area (air_aera_name)", y ="Number of air restaurants")


p2 <- air_store_info %>%
  group_by(air_genre_name) %>%
  count() %>%
  #ungroup() %>%
  top_n(15, n) %>%
  ggplot(aes(reorder(air_genre_name, n), n, fill = air_genre_name)) +
  geom_col() +
  coord_flip()+
  theme(legend.position = "none")

layout <- matrix(c(1,2), 2,1, byrow = TRUE)
multiplot(p1,p2, layout = layout)

```

## HPG Store

```{r}
leaflet(hpg_store_info) %>%
  addTiles() %>%
  addProviderTiles("cartoDB.positron") %>%
  addMarkers(~longitude, ~latitude,
             popup = ~hpg_store_id, label = ~hpg_genre_name,
             clusterOptions = markerClusterOptions()
             )
```

```{r}
p1 <- hpg_store_info %>%
  group_by(hpg_genre_name) %>% 
  count() %>%
  ggplot(aes(reorder(hpg_genre_name, n, FUN = min), n, fill = hpg_genre_name)) +
  geom_col() +
  coord_flip() +
  theme(legend.position = "none") +
  labs(x = "Type of cuisine (hpg_genre_name)", y = "Number of hpg restaurants")

p2 <- hpg_store_info %>%
  group_by(hpg_area_name) %>%
  count() %>%
  arrange(desc(n)) %>%
  head(15) %>%
  ggplot(aes(reorder(x = hpg_area_name, n, FUN = min), n, fill = hpg_area_name)) +
  geom_col() +
  coord_flip() +
  theme(legend.position = "none") +
  labs(x = "Top 15 areas (hpg_area_name)", y = "Number of hpg restaurants")

layout <- matrix(c(1,2), 1,2, byrow=TRUE)
multiplot(p1, p2, layout = layout)

```


## Holidays

```{r}
date_info %>%
  group_by(holiday_flg) %>%
  count() %>%
  ggplot(aes(x = holiday_flg, y = n, fill = holiday_flg)) +
  geom_col() +
  theme(legend.position = "none")

date_info %>%
  filter(date  > ymd("2017-04-15") & date < ymd("2017-06-01")) %>%
ggplot(aes(x = date, y= holiday_flg, color = holiday_flg)) +
geom_point() +
  theme(legend.position = "none")
```

## Test Data set

```{r}
air_visit_train <- air_visit %>%
  rename(date = visit_date) %>%
  distinct(date) %>%
  mutate(dset= "train")

bar <- sample_submission %>%
  tidyr::separate(id, c("foo","bar", "date"), sep = "_") %>%
  mutate(date = ymd(date)) %>%
  distinct(date) %>%
  mutate(dset = "test")

air_visit_train <- air_visit_train %>%
  bind_rows(bar)

year(air_visit_train$date) <- 2017

air_visit_train %>%
  filter(!is.na(date)) 
  #mutate(year = forcats::fct_relevel(as.factor(year), c("2017", "2016")))


```

# Feature relations


## Visitors per genre

```{r}
air_visit %>%
left_join(air_store_info, by = "air_store_id") %>%
  group_by(visit_date, air_genre_name) %>%
  summarise(mean_visitors = mean(visitors)) %>%
  ungroup() %>%
  ggplot(aes(visit_date, mean_visitors, color = air_genre_name)) +
  geom_line() +
  labs(y = "Average number of visitors to 'air' restaurants", x = "Date") +
  theme(legend.position = "none") +
  scale_y_log10() +
  facet_wrap(~ air_genre_name)
```


```{r}
air_visit %>%
left_join(air_store_info, by = "air_store_id") %>%
  mutate(wday = wday(visit_date, label = TRUE)) %>%
  group_by(wday, air_genre_name) %>%
 summarise(mean_visitors = mean(visitors)) %>%
  ggplot(aes(y = mean_visitors, x= air_genre_name, color = wday)) +
  geom_point() +
  theme(legend.position = "left", axis.text.y = element_blank(),
        plot.title = element_text(size = 14)) +
  coord_flip() +
  scale_color_hue()

```

```{r}
air_visit %>%
left_join(air_store_info, by = "air_store_id") %>%
  group_by(air_genre_name) %>%
  ggplot(aes(x = visitors, y = air_genre_name)) +
  ggridges::geom_density_ridges(bandwidth = 0.1) +
  scale_x_log10() +
  ggridges::scale_fill_cyclical(values = c("blue", "red"))

```


## The impact of holidays

```{r}

foo <- air_visit %>%
  #group_by(visit_date) %>%
  mutate(calendar_date = as.character(visit_date)) %>%
  left_join(date_info, by="calendar_date")

foo %>%
  ggplot(aes(holiday_flg, visitors, color = holiday_flg)) +
  geom_boxplot() +
  scale_y_log10() +
  theme(legend.position = "none")

foo %>%
  mutate(wday = wday(date, label = TRUE)) %>%
  group_by(wday, holiday_flg) %>%
  summarise(visitors_mean = mean(visitors)) %>%
  ggplot(aes(x= wday, visitors_mean, color = holiday_flg)) +
  geom_point()

```


## Restaurants per aera and the effect on visitor numbers

```{r}
air_store_info %>%
  #mutate(air_area_name = str_sub(air_area_name, 1, 12))
  group_by(air_area_name) %>%
  count(air_genre_name) %>%
  head(50) %>%
  ggplot(aes(x = air_area_name, y = air_genre_name)) +
  geom_point(aes(size = n)) +
  theme(axis.text.x = element_text(angle=45, hjust = 1, vjust= 0.9))
```


```{r}

hpg_store_info %>%
    mutate(area = str_sub(hpg_area_name, 1, 10)) %>%
  group_by(area) %>%
  count(hpg_genre_name) %>%
  head(100) %>%
  ggplot(aes(x= area, y = hpg_genre_name )) +
  geom_point(aes(size = n)) +
  theme(legend.position = "bottom",axis.text.x = element_text(angle=45, hjust = 1, vjust= 0.9))
```

```{r}
air_store_info %>%
  group_by(air_genre_name) %>%
  count(air_area_name) %>%
  ggplot(aes(reorder(x= air_genre_name, n, FUN = mean ), y = n)) +
  scale_y_log10() +
  geom_boxplot() +
  geom_jitter(color ="blue") +
  coord_flip()

```

```{r}

foo <- air_visit %>%
  left_join(air_store_info, by = "air_store_id")

bar <- air_store_info %>%
  group_by(air_genre_name, air_area_name) %>%
  count()

foobar <- hpg_store_info %>%
  group_by(hpg_genre_name, hpg_area_name) %>%
  count()

p1 <- bar %>%
  ggplot(aes(n)) +
  geom_histogram(fill = "blue", binwidth = 1) +
  labs(x = "Air genres per aera")

p2 <- foobar %>%
  ggplot(aes(n)) +
  geom_histogram(fill = "red", binwidth = 1) +
  labs(x = "HPG genres per area")


P3 <- foo %>%
  group_by(air_genre_name, air_area_name) %>%
  summarise(mean_log_visit = mean(log1p(visitors))) %>%
  left_join(bar, by = c("air_genre_name","air_area_name")) %>%
  group_by(n) %>%
  summarise(mean_mlv = mean(mean_log_visit),
          sd_mlv = sd(mean_log_visit)) %>%
  tidyr::replace_na(list(sd_mlv = 0)) %>%
  ggplot(aes(n, mean_mlv)) +
  geom_point(color = "blue", size = 4) +
  geom_errorbar(aes(ymin = mean_mlv - sd_mlv, ymax = mean_mlv + sd_mlv), width = 0.5, size = 0.7, color = "blue") +
  labs(x="Case of identifcation Air genres per area", y = "Mean +/- SD of \n mean Logp1 visitors")
```


## Reservations vs Visits

```{r}
foo <- air_reserve %>%
  mutate(visit_date = date(visit_datetime)) %>%
  group_by(air_store_id, visit_date) %>%
  summarise(reserve_visitors_air = sum(reserve_visitors))


  
  bar <- hpg_reserve %>%
    mutate(visit_date = date(visit_datetime)) %>%
    group_by(hpg_store_id, visit_date) %>%
    summarise(reserve_visitors_hpg = sum(reserve_visitors)) %>%
    inner_join(store_id_relation, by = "hpg_store_id")
  
  all_reserve <- air_visit %>%
    inner_join(foo, by = c("air_store_id", "visit_date")) %>%
    inner_join(bar, by = c("air_store_id", "visit_date")) %>%
    mutate(reserve_visitors = reserve_visitors_air + reserve_visitors_hpg)
  
  p <- all_reserve %>%
    filter(reserve_visitors < 120 ) %>%
    ggplot(aes(reserve_visitors, visitors )) +
    geom_point(color = "black", alpha = 0.5) +
    geom_abline(slope = 1, intercept = 0, color = "grey60") +
    geom_smooth(method = "lm", color = "blue") 
  
  ggMarginal(p,  type = "histogram", fill = "blue", bins=50)
    
  
```


```{r}

p1 <- all_reserve %>%
  ggplot(aes(visitors - reserve_visitors)) +
  geom_histogram(binwith = 5 , fill = "black") +
  coord_flip() +
  labs(x = "")


p2 <- all_reserve %>%
  ggplot(aes(visitors - reserve_visitors_air)) +
  geom_histogram(binwidth = 5, fill = "blue") +
  coord_flip() +
  labs(x = "")

p3 <- all_reserve %>%
  ggplot(aes(visitors - reserve_visitors_hpg)) +
  geom_histogram(binwidth = 5, fill = "red") +
  coord_flip() +
  labs(x = "")

p4 <- all_reserve %>%
  ggplot(aes(visit_date, visitors - reserve_visitors)) +
  geom_hline(yintercept = c(150, 0, -250)) +
  geom_line() +
  geom_line(aes(visit_date, visitors - reserve_visitors_air +150), color = "blue") +
  geom_line(aes(visit_date, visitors - reserve_visitors_hpg - 250), color = "red") +
  ggtitle("Visitors - Reserved: all (black), air (blue), hpg (red)")


layout <- matrix(c(4,4,2,4,4,1,4,4,3), 3, 3, byrow = TRUE)
multiplot(p1, p2, p3, p4, layout = layout)

```


```{r}
all_reserve %>%
  mutate(date = visit_date) %>%
  left_join(date_info, by = "date") %>%
  ggplot(aes(visitors - reserve_visitors, fill = holiday_flg)) +
  geom_density(alpha = 0.5)
  

```


# Feature engineering


```{r}

air_visit <- air_visit %>%
  mutate(wday = wday(visit_date, label = TRUE),
         wday = forcats::fct_relevel(wday, c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun")), month = month(visit_date, label = TRUE))

air_reserve <- air_reserve %>%
  mutate(reserve_date = date(reserve_datetime),
         reserve_hour = hour(reserve_datetime),
         reserve_wday = wday(reserve_datetime, label = TRUE),
         reserve_wday = forcats::fct_relevel(reserve_wday, c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun")),
         visit_date = date(visit_datetime),
         visit_hour = hour(visit_datetime),
         visit_wday = wday(visit_datetime, label = TRUE),
         viist_wday = forcats::fct_relevel(visit_wday, c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun")),
         diff_hour = time_length(visit_datetime - reserve_datetime, unit = "hour"),
         diff_day = time_length(visit_datetime - reserve_datetime, unit = "day"))


hpg_reserve %>%
  mutate(reserve_date = date(reserve_datetime),
         reserve_hour = hour(reserve_datetime),
         reserve_wday = wday(reserve_datetime, label = TRUE),
         reserve_wday = forcats::fct_relevel(reserve_wday, c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun")),
         visit_date = date(visit_datetime),
         visit_hour = hour(visit_datetime),
         visit_wday = wday(visit_datetime, label = TRUE),
         visit_wday = forcats::fct_relevel(visit_wday, c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun")),
         diff_hour = time_length(visit_datetime - reserve_datetime, unit = "hour"),
         diff_day = time_length(visit_datetime - reserve_datetime, unit = "day"))


# count stores in area
air_count <- air_store_info %>%
  group_by(air_area_name) %>%
  summarise(air_count = n())

hpg_count <- hpg_store_info %>%
  group_by(hpg_area_name) %>%
  summarise(hpg_count = n())

# distance
med_coord_air <- air_store_info %>%
  summarise_at(vars(longitude:latitude), median)

med_coord_hpg <- hpg_store_info %>%
  summarise_at(vars(longitude:latitude), median)



air_coords <- air_store_info %>%
  select(longitude, latitude)

hpg_coords <- hpg_store_info %>%
  select(longitude, latitude)

air_store_info$dist <- geosphere::distCosine(air_coords, med_coord_air) / 1e3
hpg_store_info$dist <- geosphere::distCosine(hpg_coords, med_coord_hpg) / 1e3

# apply counts, dist; add prefecture

air_store_info %>%
  mutate(dist_group = as.integer(case_when(
    dist < 80  ~ 1,
    dist < 300 ~ 2, 
    dist < 500 ~ 3,
    dist < 750 ~ 4,
    TRUE ~ 5 ))) %>%
  left_join(air_count, by = "air_area_name") %>%
  separate(air_area_name, c("preferture"), sep = " ", remove = FALSE)




```

## Days of the week & months of the year

```{r}
air_visit %>%
  group_by(wday) %>%
  summarise(mean_log_visitors = mean(log1p(visitors)),
            sd_log_visitors = sd(log1p(visitors))) %>%
  ggplot(aes(wday, mean_log_visitors, color = wday)) +
  geom_point(size = 4) +
  geom_errorbar(aes(ymin = mean_log_visitors - sd_log_visitors,
                    ymax = mean_log_visitors + sd_log_visitors,
                    color = wday), width = 0.5, size = 0.7 ) +
  theme(legend.position = "none")


p2 <- air_visit %>%
  mutate(visitors = log1p(visitors)) %>%
  ggplot(aes(visitors, wday, fill = wday)) +
  ggridges::geom_density_ridges(bandwidth = 0.1) +
  scale_x_log10() +
  theme(legend.position = "none") +
  labs(x= "log1P(visitors", y ="")

p3 <- air_visit %>%
  group_by(month) %>%
  summarise(mean_log_visitors = mean(log1p(visitors)),
            sd_log_visitors = sd(log1p(visitors))) %>%
  ggplot(aes(month, mean_log_visitors, color = month)) +
  geom_point(size = 4) +
  geom_errorbar(aes(ymin = mean_log_visitors - sd_log_visitors,
                    ymax = mean_log_visitors + sd_log_visitors,
                    color = month), width = 0.5, size = 0.7) +
  theme(legend.position = "none")

p4 <- air_visit %>%
  mutate(visitors = log1p(visitors)) %>%
  ggplot(aes(visitors, month, fill = month)) +
  geom_density_ridges(bandwidth = 0.1) +
  scale_x_log10() +
  theme(legend.position = "none") +
  labs(x = "log1p(visitors)", y = "")

layout <- matrix(c(1,2,3,4),2,2,byrow=TRUE)

multiplot(p1, p2, p3, p4, layout=layout)
```



# Time series parameters

```{r}



air_visit %>%
  left_join(air_store_info, by = "air_store_id") %>%
  group_by(air_store_id, air_genre_name) %>%
  summarise(mean_log_visits = mean(log1p(visitors)),
            mean_log_visits = mean(log1p(visitors)),
            sd_log_visits = sd(log1p(visitors))) %>%
  ungroup()

params_ts1 <- function(rownr){
  
  bar <- air_visit %>%
    filter(air_store_id == foo$air_store_id[rownr])
  slope <- summary(lm(visitors ~ visit_date, data = bar))$coef[2]
  slope_err <- summary(lm(visitors ~ visit_date, data = bar))$coef[4]
 
  foobar <- tibble(
    air_store_id = foo$air_store_id[rownr],
    slope = slope,
    slope_err = slope_err
  )
  return(foobar)

}

params <- params_ts1(1)

for (i in seq(2,nrow(foo))){

  params <- bind_rows(params, params_ts1(i))

}



ts_params <- foo %>%
  left_join(params, by = "air_store_id")
```




