---
title: "Reticulate: Set Python environment"
author: "Karim Mezhoud"
date: '`r Sys.Date()`'
output:
  html_document:
    number_sections: true
    fig_caption: true
    toc: true
    fig_width: 7
    fig_height: 6
    theme: cosmo
    highlight: tango
    code_folding: show #hide
    self_contained: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, error=TRUE, warning =TRUE)
knitr::opts_chunk$set(fig.width=12, fig.height=8) 
```



# Load packages
```{r}
library(reticulate)
```

# Locating python, anaconda and environments

## bash Chunk
```{bash}
# which last installed version from bash
python3 --version
python2 --version
```

```{bash}
# where is executable python
python -c "import sys; print(sys.executable)"
python2.7 -c "import sys; print(sys.executable)"
python3.7 -c "import sys; print(sys.executable)"

```

```{bash}
# where is installed each python version
which -a python2.7 python3 python
```

## `R` Chunk

```{r}
# which version used by R
scrobbler::py_version()
```

To figure out where your python 3 is installed type the following into the R console

### Where python is installed
```{r}
# where python is installed
base::Sys.which("python")

# where python2 is installed
base::Sys.which("python2")

# where python3 is installed
base::Sys.which("python3")
```




# Setting python version

If your version has a 2 as its major number then you’ll need to tell R the path of the version 3. This requires editing your `.Rprofile` file to set python 3 as your default.

### Edit r profile
```{r}
usethis::edit_r_profile()
```
and add something like this

```{r}
Sys.setenv(RETICULATE_PYTHON = "/usr/local/bin/python2")
```

# `Reticulate` setting
Each version of Python on the system has it’s own set of packages. Reticulate will automatically find a version of Python that contains the first package that you import from R.

`reticulate` uses default setting of r-session or r-profile. The default path for the first installation is `/usr/bin/python`. 
It is commun that multiple versions of python are installed in the same machine. The second and next version will be installed to `/usr/local/bin/python3`. for example:


* `Anaconda` distribution uses self environment where install all needed packages. We can found also python into anaconda environment path `/Users/Mezhoud/miniconda3/envs/Ana_Env/bin/python`.


* `r-tensorflow` uses also self environment and uses self python and package version `/Users/Mezhoud/.virtualenvs/r-tensorflow/bin/python`. 


## check configuration

```{r}
reticulate::py_config()
```
The first part shows what R session is using:
* wich version: python2.7.10
* Path of libraries
* Path of the home
* numpy version and path

The second part shows the others installed version. we can see `python3, python, miniconda` distribution and python used by `R-tensorflow`.
This command is usefull. It screen existing setting and installed versions of python.

A second function that display the which used and available python verisons.
```{r}
reticulate::py_discover_config()
```

We can look for anaconda distribution.

```{r}
# where Anaconda is installed
reticulate::conda_list()

```

We can look for packages in used python 

```{r}
reticulate::py_module_available("pandas")
```

## change python version
The default setting remain the best configuration for reticulate. This [link](https://github.com/rstudio/reticulate/issues/368#issuecomment-487648765) shows some non-concordance between versions, installation ways (pip, conda, ), and environments ([Pipenv vs virtualenv vs conda environment](https://medium.com/@krishnaregmi/pipenv-vs-virtualenv-vs-conda-environment-3dde3f6869ed)).


```{r}
## we have to refresh R session before to set a new python version
.rs.restartR()
reticulate::use_python("/Users/Mezhoud/miniconda3/envs/Ana_Env/bin/python", required = TRUE)

## not necessary if we set use_python to Ana_Env
#reticulate::use_condaenv( condaenv = "/Users/Mezhoud/miniconda3/envs/Ana_Env/bin/python", conda = "auto", required = TRUE)
```


```{r}
reticulate::py_config()
```

```{r}
reticulate::py_module_available("matplotlib")
```

```{r}
# which version used by R
scrobbler::py_version()
```


# python code

```{r}
# refresh session and use the default python path
.rs.restartR()
```

```{python}
import numpy as np
import matplotlib.pyplot as plt

t = np.arange(0.0, 2.0, 0.01 )
s = 1 + np.sin(2*np.pi*t)
plt.plot(t,s)
```






