---
title: "Predict the number of attempts to get the right answer"
date: '`r Sys.Date()`'
output:
  html_document:
    number_sections: true
    fig_caption: true
    toc: true
    fig_width: 7
    fig_height: 4.5
    theme: cosmo
    highlight: tango
    code_folding: show #hide
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo=TRUE, error=FALSE)
```

# Introduction
- Evaluation: the quadratic weighted kappa 
- Apps: Bird Measurer, Cart Balancer, Cauldron Filler, Chest Sorter, and Mushroom Sorter.
- The outcomes in this competition are grouped into 4 groups (labeled accuracy_group in the data): 0, 1, 2, 3.

## Load librairies

```{r include=FALSE, echo=TRUE}
library(stringr) # string manipulation
library(lubridate)
library(ggplot2)
library(tidyr)
library(tidyverse)
library(data.table)
library(jsonlite)
library(gridExtra)
```

# Explore and understand the data

```{r}
train <- read_rds("datasets/train.rds")
train_labels <- read_rds("datasets/train_labes.rds")
test <- read_rds("datasets/test.rds")
specs <- read_rds("datasets/specs.rds")
sample_submission <- read_rds("datasets/sample_submission.rds")
```

## reduce the size by title (game)  and by event_code wich targets only assessement attempts

* The assessment is done only for 5 titles: we filter by these games
* The attempts are indicated by event_code: 4100 and 4100. We filter by these two event_codes.

```{r}
get_accuracy <- function(df, game ,code){
tmp <- df %>%
  filter(str_detect(title , pattern = game)) %>%
  filter(event_code ==  code) %>%
  mutate(event_data = gsub('"', '', event_data)) %>%
  mutate(Status =  ifelse(str_detect(event_data, pattern = 'correct:true'),  'correct',
                    ifelse(str_detect(event_data, pattern = 'correct:false'),  'incorrect',
                           NA))) %>%
  dplyr::group_by(installation_id, game_session, title) %>%
  summarise(num_correct = sum(Status %in% 'correct'),
            num_incorrect = sum(Status %in% 'incorrect')) %>%
  ungroup %>%
  mutate(accuracy = num_correct/(num_correct+num_incorrect)) %>%
  mutate(accuracy_group = ifelse(accuracy == 0,  0L,
          ifelse(accuracy> 0 & accuracy < 0.5,  1L,
          ifelse(accuracy >= 0.5 & accuracy < 1,  2L,
                  3L
          )))) 
  return(tmp)
}


bird_accuracy <- get_accuracy(train ,'Bird Measurer' , '4110')
cart_accuracy <- get_accuracy(train, "Cart Balancer", '4100')
caudron_accuracy <- get_accuracy(train, "Cauldron Filler", "4100")
chest_accuracy <- get_accuracy(train, "Chest Sorter", "4100")
mushroom_accuracy <- get_accuracy(train, "Mushroom Sorter", "4100")

train_labels_own <- dplyr::bind_rows(bird_accuracy, cart_accuracy, caudron_accuracy, chest_accuracy, mushroom_accuracy) 

head(train_labels_own)


paste0("Is dimensions (x,y) of train_labels and trains_labels_own are equal:", dplyr::all_equal(train_labels, train_labels_own))
```

```{r}
right_join(train_labels, train_labels_own, by = c("game_session", "installation_id")) %>%
  mutate(match = ifelse(accuracy_group.x == accuracy_group.y, TRUE, FALSE)) %>% select(match) %>% table
```

**Good!**

```{r}
# function to convert datetime and get time fetaures
convert_timestamp <- function(df){
  tmp <- df %>%
  mutate(timestamp = gsub( "T", " ", timestamp)) %>%
  mutate(timestamp = gsub( "\\..*Z$", "", timestamp)) %>%
  mutate(Date = as_datetime(format(as.POSIXct(timestamp ,format="%Y-%m-%d %H:%M:%S"),"%Y-%m-%d %H:%M:%S"))) %>%
  mutate(game_time = format( as.POSIXct(Date)+game_time/1000, "%Y-%m-%d %H:%M:%S")) %>%
  arrange(Date) %>%
  mutate(timing = ymd_hms(game_time) - ymd_hms(Date)) #%>% 
  #select(Date, installation_id, game_session, event_id, game_time, event_count, timing,  event_code, title) 
  
  return(tmp)
}

```



# Compute test_labels

```{r}


bird_accuracy_ <- get_accuracy(test ,'Bird Measurer' , '4110')
cart_accuracy_ <- get_accuracy(test, "Cart Balancer", '4100')
caudron_accuracy_ <- get_accuracy(test, "Cauldron Filler", "4100")
chest_accuracy_ <- get_accuracy(test, "Chest Sorter", "4100")
mushroom_accuracy_ <- get_accuracy(test, "Mushroom Sorter", "4100")

test_labels <- dplyr::bind_rows(bird_accuracy_, cart_accuracy_, caudron_accuracy_, chest_accuracy_, mushroom_accuracy_) 

class(bird_accuracy)

test_labels %>% head(1000) %>% DT::datatable()

```

* There is multiple attempts with assessment for each instllation_id, game_session.
* We support that the last assessment is the final result.
* We need to extract le last assessment by the datetime variable
* We need to merge test_labels with original label to extract datetime.

## Associate test_labels (accuracies) to original test data 


```{r}
# Convert datatime
test_ <- convert_timestamp(test)


## get the last assessment for each game_session
 sub_test_labels_1 <- test_ %>%
  left_join(test_labels, by = c("installation_id", "game_session", "title")) %>%
  filter(!is.na(accuracy_group)) %>% 
  arrange(desc(timestamp)) %>% 
  distinct(installation_id, .keep_all = TRUE) 
 
  #group_by(installation_id) %>%  as.data.frame() %>% 
  #slice(which.max(Date)) %>%
  #ungroup() %>%
  #select(installation_id, accuracy_group) %>%
  #right_join(sample_submission, by = "installation_id") %>%
  #filter(!is.na(accuracy_group.x))  %>%
  # select(installation_id, accuracy_group.x) %>%
   #rename(accuracy_group = accuracy_group.x)
 
 sub_test_labels_1 %>% head(557) %>% DT::datatable()
 
```

# Look for information of missed instllation_id accuracies

## Get accuracy from event data

We will ignore event_code 4100 and 4110 and try to extract if is there string like `correct:true` & `correct:false`.

```{r}
 missed_install <- test_ %>%
 anti_join( sub_test_labels_1 , by = c("installation_id")) %>%
  mutate(event_data = gsub('"', '', event_data)) %>%
  mutate(Status =  ifelse(str_detect(event_data, pattern = 'correct:true'),  'correct',
                    ifelse(str_detect(event_data, pattern = 'correct:false'),  'incorrect',
                           NA)))

 test_last_assessment  <- missed_install  %>% 
                        filter(type == "Assessment") %>% 
                       arrange(desc(timestamp)) %>% 
                       distinct(installation_id, .keep_all = TRUE)


sub_test_labels_2 <- test_last_assessment%>%
  filter(!is.na(Status)) %>%
    group_by(installation_id, game_session, title) %>%
  summarise(num_correct = sum(Status %in% 'correct'),
            num_incorrect = sum(Status %in% 'incorrect'),
            num_NA = sum(is.na(Status))) %>%
 # ungroup %>%
  mutate(accuracy = num_correct/(num_correct+num_incorrect+num_NA)) %>%
  mutate(accuracy_group = ifelse(accuracy == 0,  0L,
          ifelse(accuracy> 0 & accuracy < 0.5,  1L,
          ifelse(accuracy >= 0.5 & accuracy < 1,  2L,
                  3L
          )))) %>%
  left_join(test_, by = c("installation_id", "game_session", "title")) %>%
  #filter(!is.na(accuracy_group)) %>%
  group_by(installation_id) %>%
  slice(which.max(Date)) %>%
  select(installation_id, accuracy_group) %>%
  right_join(sample_submission, by = "installation_id") %>%
  filter(!is.na(accuracy_group.x))  %>%
  select(installation_id, accuracy_group.x) %>%
  rename(accuracy_group = accuracy_group.x)

sub_test_labels_2 %>% head(291) %>% DT::datatable()
```


# Merge  sub_test_labels_1 and sub_test_labels_2

```{r}
sub_test_labels_848 <- 
  bind_rows(sub_test_labels_1, sub_test_labels_2) %>%
  right_join(sample_submission, by = "installation_id") %>%
  filter(!is.na(accuracy_group.x)) %>%
  select(installation_id, accuracy_group.x) %>%
  rename(accuracy_group = accuracy_group.x)

sub_test_labels_848 %>% head(848) %>% DT::datatable()
```

## The second rest of installation_id that we did not find any information in event_data

* The remain assessment have a timing 0 secs. 
* we will try to find installation_id from train data that have the same proprieties, like `game_time`, `title`, `datetime`.

```{r}
train_useful <- train %>%
  filter(type == "Assessment") %>%
  filter(event_code == 4100| event_code == 4110)
train_useful %>% head
```

## Here is the proprieties of the second rest of instllation_id without accuracy

There are 163 instllation_id mainly with game_time equal to 0 except `Chest Sorter (Assessment)`.

```{r}
  bind_rows(sub_test_labels_1, sub_test_labels_2) %>%
  right_join(sample_submission, by = "installation_id") %>%
  filter(is.na(accuracy_group.x)) %>%
  select(installation_id, accuracy_group.x) %>%
  rename(accuracy_group = accuracy_group.x) %>%
  left_join(test, by = c('installation_id')) %>%
  ungroup() %>%
  filter(type == "Assessment") #%>%
  select( "game_time", "title", "world") %>%
  left_join(train_useful, by= c(  "game_time", "title")) %>%
    DT::datatable()
```

* In the next code we focus on   `event_count`, `title`, `world` to find common rows with train_useful data. 


```{r}
bind_rows(sub_test_labels_1, sub_test_labels_2) %>%
  right_join(sample_submission, by = "installation_id") %>%
  filter(is.na(accuracy_group.x)) %>%
  select(installation_id, accuracy_group.x) %>%
  rename(accuracy_group = accuracy_group.x) %>%
  left_join(test, by = c('installation_id')) %>%
  ungroup() %>%
  filter(type == "Assessment") %>%
  select("installation_id", "event_count","game_time", "title", "world") %>%
  inner_join(train_useful, by= c( "event_count","title", "world")) #%>%  #DT::datatable()
  
```

* We find 29 instllation_id from train that have the same `event_count`, `title`, `world`. We will use its accuracies

```{r}
 sub_test_labels_3 <- bind_rows(sub_test_labels_1, sub_test_labels_2) %>%
  right_join(sample_submission, by = "installation_id") %>%
  filter(is.na(accuracy_group.x)) %>%
  select(installation_id, accuracy_group.x) %>%
  rename(accuracy_group = accuracy_group.x) %>%
  left_join(test, by = c('installation_id')) %>%
  ungroup() %>%
  filter(type == "Assessment") %>%
  select( "installation_id","event_count","game_time", "title", "world") %>%
  inner_join(train_useful, by= c( "event_count","title", "world")) %>%
  select(installation_id.x,installation_id.y,  game_session, title) %>% 
  rename(installation_id = installation_id.y) %>%
  left_join(train_labels, by = c( "installation_id", "game_session", "title")) %>%
  group_by(installation_id.x, title) %>%
  summarise(new_accuracy_group = median(accuracy_group)) %>%
  rename(accuracy_group = new_accuracy_group,
         installation_id = installation_id.x) %>%
  select(installation_id, accuracy_group)

sub_test_labels_3 %>% DT::datatable()
```

#For the moment we fill NA by 0
```{r}
submission <- bind_rows(sub_test_labels_1, sub_test_labels_2, sub_test_labels_3) %>%
  right_join(sample_submission, by = "installation_id") %>%
  #filter(is.na(accuracy_group.x)) %>%
  select(installation_id, accuracy_group.x) %>%
  rename(accuracy_group = accuracy_group.x) %>%
  mutate(accuracy_group = ifelse(is.na(accuracy_group), 0, accuracy_group))

fwrite(submission, file = "submission.csv")

submission$accuracy_group %>% table() %>% names()
```


