---
title: "testing and EDA"
date: '`r Sys.Date()`'
output:
  html_document:
    number_sections: true
    fig_caption: true
    toc: true
    fig_width: 7
    fig_height: 4.5
    theme: cosmo
    highlight: tango
    code_folding: show #hide
--- 

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo=TRUE, error=FALSE)
```

# Introduction
- Evaluation: the quadratic weighted kappa 
- Apps: Bird Measurer, Cart Balancer, Cauldron Filler, Chest Sorter, and Mushroom Sorter.
- The outcomes in this competition are grouped into 4 groups (labeled accuracy_group in the data): 0, 1, 2, 3.

## Load librairies

```{r include=FALSE, echo=TRUE}
library(stringr) # string manipulation
library(lubridate)
library(ggplot2)
library(tidyr)
library(tidyverse)
library(data.table)
library(jsonlite)
library(gridExtra)
```

## Reduce the size of files

```{r}
## reduce file size from csv to rds format
csv2rds <- function(path){
  
  for(f in seq(list.files(path))){
    
    tmp <- fread(list.files(path, full.names = TRUE)[f])
    fname <-  gsub("\\.csv", "",  list.files(path, pattern="\\.csv$")[f])
    saveRDS(tmp, paste0(path,"/", fname, ".rds", sep=""))
    rm(tmp)
  }
  
  
}

#csv2rds("datasets")

  # train_labels <- fread("datasets/train_labels.csv", stringsAsFactors = FALSE)
  # train <- fread("datasets/train.csv", stringsAsFactors = TRUE)
  # test <- fread("datasets/test.csv", stringsAsFactors = FALSE)
  # sample_submission <- fread("datasets/sample_submission.csv", stringsAsFactors = FALSE)
  # specs <- fread("datasets/specs.csv", stringsAsFactors = TRUE)

# saveRDS(train_labels, file = "datasets/train_labes.rds")
#saveRDS(train, file= "datasets/train.rds")
# saveRDS(test, file = "datasets/test.rds")
# saveRDS(sample_submission, file = "datasets/sample_submission.rds")
# saveRDS(specs, file = "datasets/specs.rds")
# 
# rm(train, train_labels, test, sample_submission, specs)
# 
# gc()

```

## Load files


# Where is assessment result per istallation_id?


```{r}
get_accuracy <- function(df, game ,code){
tmp <- df %>%
  filter(str_detect(title , pattern = game)) %>%
  filter(event_code == code) %>%
  # mutate(timestamp = gsub( "T", " ", timestamp)) %>%
  # mutate(timestamp = gsub( "\\..*Z$", "", timestamp)) %>%
  # mutate(Date = as_datetime(format(as.POSIXct(timestamp ,format="%Y-%m-%d %H:%M:%S"),"%Y-%m-%d %H:%M:%S"))) %>%
  # mutate(game_time = format( as.POSIXct(Date)+game_time/1000, "%Y-%m-%d %H:%M:%S")) %>%
  # arrange(Date) %>%
  mutate(event_data = gsub('"', '', event_data)) %>%
  mutate(Status =  ifelse(str_detect(event_data, pattern = 'correct:true'),  'correct',
                    ifelse(str_detect(event_data, pattern = 'correct:false'),  'incorrect',
                           NA))) %>%
 # mutate(timing = ymd_hms(game_time) - ymd_hms(Date)) %>%
  group_by(installation_id, game_session, title) %>%
  summarise(num_correct = sum(Status %in% 'correct'),
            num_incorrect = sum(Status %in% 'incorrect'),
            num_NA = sum(is.na(Status))) %>%
  ungroup %>%
  mutate(accuracy = num_correct/(num_correct+num_incorrect+num_NA)) %>%
  mutate(accuracy_group = ifelse(accuracy == 0,  0,
          ifelse(accuracy> 0 & accuracy < 0.5,  1,
          ifelse(accuracy >= 0.5 & accuracy < 1,  2,
                  3
          )))) 
  return(tmp)
}


bird_accuracy <- get_accuracy(train ,'Bird Measurer' , '4110')
cart_accuracy <- get_accuracy(train, "Cart Balancer", '4100')
caudron_accuracy <- get_accuracy(train, "Cauldron Filler", "4100")
chest_accuracy <- get_accuracy(train, "Chest Sorter", "4100")
mushroom_accuracy <- get_accuracy(train, "Mushroom Sorter", "4100")


```

## There is non common pairs(`installation_id, game_session`), we can rbind accuracies

```{r}
train_labels_own <- dplyr::bind_rows(bird_accuracy, cart_accuracy, caudron_accuracy, chest_accuracy, mushroom_accuracy) #%>%
  #arrange(installation_id) #%>% head(1000) %>% DT::datatable()
```

* The accuracies were done for each game
* One we compute accuracies for the 5 games

# 557 accuracies of of instllation_id is in event_data
```{r}
dplyr::bind_rows(bird_accuracy, cart_accuracy, caudron_accuracy, chest_accuracy, mushroom_accuracy) %>% #select(installation_id) %>% n_distinct()
   group_by(installation_id) %>%
  summarise(num_correct_all = sum(num_correct),
            num_incorrect_all = sum(num_incorrect),
            num_NA_all = sum(is.na(num_NA))) %>%
  ungroup %>%
  mutate(accuracy_all = num_correct_all/(num_correct_all+num_incorrect_all+num_NA_all)) %>%
  mutate(accuracy_group_all = ifelse(accuracy_all == 0,  0,
          ifelse(accuracy_all > 0 & accuracy_all < 0.5,  1,
          ifelse(accuracy_all >= 0.5 & accuracy_all < 1,  2,
                  3
          ))))  
```


# Timing Feature engineer

```{r}
convert_timestamp <- function(df, game, code){
  tmp <- df %>%
  filter(str_detect(title , pattern = game)) %>%
  filter(event_code %in% code) %>%
  mutate(timestamp = gsub( "T", " ", timestamp)) %>%
  mutate(timestamp = gsub( "\\..*Z$", "", timestamp)) %>%
  mutate(Date = as_datetime(format(as.POSIXct(timestamp ,format="%Y-%m-%d %H:%M:%S"),"%Y-%m-%d %H:%M:%S"))) %>%
  mutate(game_time = format( as.POSIXct(Date)+game_time/1000, "%Y-%m-%d %H:%M:%S")) %>%
  arrange(Date) %>%
    #mutate(event_data = gsub('"', '', event_data)) %>%
  #mutate(Status =  ifelse(str_detect(event_data, pattern = 'correct:true'),  'correct',
   #                 ifelse(str_detect(event_data, pattern = 'correct:false'),  'incorrect',
    #                       NA))) %>%
  mutate(timing = ymd_hms(game_time) - ymd_hms(Date)) %>% 
  select(Date, installation_id, game_session, event_id, game_time, event_count, timing,  event_code, title) #%>% filter(timing > 1000)
  
  return(tmp)
}

  convert_timestamp(train , c("Bird Measurer","Cart Balancer", "Mushroom Sorter") , c("4110", "4100"))


```

* The correct attempt is always the last one before to process to the next game_session. Or there is no correct answer.

## How long time has the game session to get the correct answer?

```{r}
convert_timestamp(train ,'Bird Measurer' , '4110') %>%
  group_by(installation_id, game_session) %>%
  summarise(longtime = sum(timing)) %>%
  right_join(convert_timestamp(train ,'Bird Measurer' , '4110') , by= c("installation_id", "game_session"))
```

## `Mean`, `median` and `sd` and timing
```{r}
convert_timestamp(train ,'Bird Measurer' , '4110') %>%
  group_by(installation_id) %>%
  summarise(mean_timing = mean(timing),
            median_timing = median(timing),
            sd_timing = sd(timing))
```


# Event count and attempts featuring

## How many attempts per instllation_id to get the succes the game_session?
we need to group by instllation_id and count `Status` column.

```{r}
  convert_timestamp(train ,'Bird Measurer' , '4110') %>%
  group_by(installation_id) %>%
  summarise(attempts = n()) %>%
  right_join(convert_timestamp(train ,'Bird Measurer' , '4110') , by = "installation_id")
```


## mean, median, sd of event_count
```{r}
convert_timestamp(train ,'Bird Measurer' , '4110') %>%
  group_by(installation_id) %>%
  summarise(mean_event_count = mean(event_count),
            median_event_count = median(mean_event_count),
            sd_event_count = sd(event_count,na.rm = TRUE))
```

# Datetime , months, week, day features
```{r}
convert_timestamp(train ,'Bird Measurer' , '4110') %>%
mutate(datetime = as.numeric(as.POSIXct(Date ,format="%Y-%m-%d %H:%M:%S"))) %>%
mutate(Date = as.Date(Date)) %>%
mutate(Months = lubridate::month(ymd(Date))) %>%
mutate(Week = lubridate::week(ymd(Date))) %>%
mutate(wday = lubridate::wday(ymd(Date)))
```


```{r}
train %>%
left_join(train_labels, by = c("game_session", "installation_id", "title")) %>%
  filter(!is.na(accuracy_group)) %>%
  group_by(accuracy_group) %>%
  summarise(Frequencies = n()) %>%
  ggplot() +
  aes(x = accuracy_group, y = Frequencies) +
  geom_col()+
  #theme( axis.text.x  = element_text(angle=45, hjust=1, vjust=0.9)) +
  labs(title = paste0("Accuracy distribution of train_labels data"))
```

```{r}
train %>%
left_join(train_labels, by = c("game_session", "installation_id", "title")) %>%
  filter(accuracy_group == 3) %>%
  mutate(event_data = gsub('"', '', event_data)) %>%
    mutate(Correct =  ifelse(str_detect(event_data, pattern = 'correct:true'),  'correct',
                      ifelse(str_detect(event_data, pattern = 'correct:false'),  'false',
                             NA))) %>%
  filter(game_time == 0) %>%
  select(installation_id, title, type, game_session, game_time,Correct, num_correct, num_incorrect, accuracy, accuracy_group)
  #filter(!is.na(Correct)) %>%
  #select(installation_id, title, type, game_session, game_time, event_id, event_count, Correct ,everything()) %>%
  group_by(installation_id) %>%
  summarise(Num_correct = sum(Correct %in% 'correct'),
            Num_incorrect = sum(Correct %in% 'false'),
            num_NA = sum(is.na(Correct))) %>%
  ungroup %>%
  mutate(Accuracy = Num_correct/(Num_correct+Num_incorrect+num_NA)) %>%
  mutate(Accuracy_group = ifelse(Accuracy == 0,  0,
          ifelse(Accuracy> 0 & Accuracy < 0.5,  1,
          ifelse(Accuracy >= 0.5 & Accuracy < 1,  2,
                  3
          )))) %>%
  group_by(Accuracy_group) %>%
  summarise(Frequencies = n()) %>%
  ggplot() +
  aes(x = Accuracy_group, y = Frequencies) +
  geom_col()+
  labs(title = paste0("Manual Accuracy distribution train/train_labels data"))
```
# Split Date and Time

```{r}
train <- train %>% 
mutate(timestamp = gsub( "T", " ", timestamp)) %>%
mutate(timestamp = gsub( "\\..*Z$", "", timestamp)) %>%
mutate(Date = format(as.POSIXct(timestamp ,format="%Y-%m-%d %H:%M:%S"),"%Y-%m-%d")) %>%
mutate(Months = lubridate::month(ymd(Date), label =TRUE)) %>%
mutate(Time = format(as.POSIXct(timestamp ,format="%Y-%m-%d %H:%M:%S"),"%H:%M")) %>%
mutate(wday = lubridate::wday(ymd(Date), label = TRUE)) %>%
select(Date, wday, Time, everything(), -timestamp)
  
train %>% head
```


```{r, fig.height= 6}
p1 <- test %>%
  filter(event_code == 4100| event_code == 4110)%>% 
    mutate(event_data = gsub('"', '', event_data)) %>%
    mutate(Correct =  ifelse(str_detect(event_data, pattern = 'correct:true'),  'correct',
                      ifelse(str_detect(event_data, pattern = 'correct:false'),  'false',
                             NA))) %>%
  #filter(!is.na(Correct)) %>%
  group_by(installation_id) %>%
  summarise(num_correct = sum(Correct %in% 'correct'),
            num_incorrect = sum(Correct %in% 'false'),
            num_NA = sum(is.na(Correct))) %>%
    ungroup %>%
  mutate(accuracy = num_correct/(num_correct+num_incorrect + num_NA)) %>%
  mutate(accuracy_group = ifelse(accuracy == 0,  0,
          ifelse(accuracy> 0 & accuracy < 0.5,  1,
          ifelse(accuracy >= 0.5 & accuracy < 1,  2,
                  3
          )))) %>%
  group_by(accuracy_group) %>%
  summarise(Frequencies = n()) %>%
  ggplot() +
  aes(x = accuracy_group, y = Frequencies) +
  geom_col()+
  #theme( axis.text.x  = element_text(angle=45, hjust=1, vjust=0.9)) +
  labs(title = paste0("Accuracy distribution of test data: event_code == 4100|4110"))


p2 <- test %>%
  #filter(event_code == 4100| event_code == 4110)%>% 
    mutate(event_data = gsub('"', '', event_data)) %>%
    mutate(Correct =  ifelse(str_detect(event_data, pattern = 'correct:true'),  'correct',
                      ifelse(str_detect(event_data, pattern = 'correct:false'),  'false',
                             NA))) %>%
  filter(!is.na(Correct)) %>%
  group_by(installation_id) %>%
  summarise(num_correct = sum(Correct %in% 'correct'),
            num_incorrect = sum(Correct %in% 'false'),
            num_NA = sum(is.na(Correct))) %>%
    ungroup %>%
  mutate(accuracy = num_correct/(num_correct+num_incorrect+ num_NA)) %>%
  mutate(accuracy_group = ifelse(accuracy == 0,  0,
          ifelse(accuracy> 0 & accuracy < 0.5,  1,
          ifelse(accuracy >= 0.5 & accuracy < 1,  2,
                  3
          )))) %>%
  group_by(accuracy_group) %>%
  summarise(Frequencies = n()) %>%
  ggplot() +
  aes(x = accuracy_group, y = Frequencies) +
  geom_col()+
  #theme( axis.text.x  = element_text(angle=45, hjust=1, vjust=0.9)) +
  labs(title = paste0("Accuracy distribution of test data from event_data: !is.na(Correct)"))

gridExtra::grid.arrange(p1, p2, nrow = 2)

```


```{r}
sub_582 <- test %>%
  filter(event_code == 4100| event_code == 4110)%>% 
    mutate(event_data = gsub('"', '', event_data)) %>%
    mutate(Correct =  ifelse(str_detect(event_data, pattern = 'correct:true'),  'correct',
                      ifelse(str_detect(event_data, pattern = 'correct:false'),  'false',
                             NA))) %>%
  #filter(!is.na(Correct)) %>%
  group_by(installation_id) %>%
  summarise(num_correct = sum(Correct %in% 'correct'),
            num_incorrect = sum(Correct %in% 'false'),
            num_NA = sum(is.na(Correct))) %>%
    ungroup %>%
  mutate(accuracy = num_correct/(num_correct+num_incorrect + num_NA)) %>%
  mutate(accuracy_group = ifelse(accuracy == 0,  0,
          ifelse(accuracy> 0 & accuracy < 0.5,  1,
          ifelse(accuracy >= 0.5 & accuracy < 1,  2,
                  3
          )))) 


sub_848 <- test %>%
  #filter(event_code == 4100| event_code == 4110)%>% #select(installation_id) %>% n_distinct()
    mutate(event_data = gsub('"', '', event_data)) %>%
    mutate(Correct =  ifelse(str_detect(event_data, pattern = 'correct:true'),  'correct',
                      ifelse(str_detect(event_data, pattern = 'correct:false'),  'false',
                             NA))) %>%
  filter(!is.na(Correct)) %>%
  group_by(installation_id) %>%
  summarise(num_correct = sum(Correct %in% 'correct'),
            num_incorrect = sum(Correct %in% 'false'),
            num_NA = sum(is.na(Correct))) %>%
  ungroup %>%
  mutate(accuracy = num_correct/(num_correct+num_incorrect+ num_NA)) %>%
  mutate(accuracy_group = ifelse(accuracy == 0,  0,
          ifelse(accuracy> 0 & accuracy < 0.5,  1,
          ifelse(accuracy >= 0.5 & accuracy < 1,  2,
                  3
          ))))
  
  
  
missed_installation_id_for_848 <- sample_submission %>%
  left_join(sub_848, by= "installation_id") %>%
  filter(is.na(accuracy_group.y)) %>%
  select(installation_id)
 

missed_installation_id_for_582 <- sample_submission %>%
  left_join(sub_582, by= "installation_id") %>%
  filter(is.na(accuracy_group.y)) %>%
  select(installation_id)

test %>%
  filter(installation_id %in% missed_installation_id_for_848$installation_id) %>% 
  mutate(event_data = gsub('"', '', event_data)) %>%
  #filter(type == "Clip") %>%
  left_join(specs, by = "event_id") %>%
  mutate(args = gsub('"', '', args)) %>%
   mutate(Correct   =  ifelse(str_detect(info, pattern = 'correct'),  'Correct', NA),
          Incorrect = ifelse(str_detect(info, pattern = 'Incorrect'), 'Incorrect',NA)) %>%
    select(event_id,Correct, Incorrect, event_data, args, info)
```

```{r}
specs %>%
  #filter(event_id == "3bfd1a65") %>% 
  mutate(args = gsub('"', '', args)) %>%
  mutate(info = gsub('"', '', info)) %>%
  mutate(Correct =  ifelse(str_detect(info, pattern = 'Correct'),  'correct',
                      ifelse(str_detect(info, pattern = 'false'),  'false',
                             NA))) 
```
```{r}
train_labels
```


```{r}
 sample_submission %>%
left_join(sub_848, by = 'installation_id') %>%
mutate(accuracy_group.y = ifelse(is.na(accuracy_group.y), 0, accuracy_group.y)) %>%
select(installation_id, accuracy_group.y) %>%
rename( accuracy_group = accuracy_group.y)
```

```{r}
specs %>%
  filter(event_id == "27253bdc")
```

```{r}
test %>%
  filter(installation_id == "00abaee7") %>%
    mutate(event_data = gsub('"', '', event_data)) %>%
    mutate(Correct =  ifelse(str_detect(event_data, pattern = 'correct:true'),  'correct',
                      ifelse(str_detect(event_data, pattern = 'correct:false'),  'false',
                             NA)))
  
```

```{r}
sample_submission %>%
  filter(installation_id == "00abaee7")
```


```{r}
train_labels %>%
  filter(installation_id == "00abaee7")
```

# How many game_session per installation_id?
```{r, fig.width= 12}
train %>% #filter(event_code == 4100) %>% select(installation_id) %>% unique
  filter(installation_id == "0006a69f") %>%
  mutate(event_data = gsub('"', '', event_data)) %>%
  mutate(Correct = str_detect(event_data, pattern = 'correct:true')) %>%
  group_by(game_session, Correct ) %>%
  summarise(`Duration (Min)` = log1p(sum(game_time/60000))) %>%
  ggplot()+
  aes(x = game_session, y = `Duration (Min)`, fill = Correct) +
  geom_col()+
    theme( axis.text.x  = element_text(angle=45, hjust=1, vjust=0.9)) +
  labs(title = paste0("installation_id:",  "0001e90f"))
  
```


```{r}
glimpse(train_labels)
```

```{r}
glimpse(specs)
```

There are `installation_id` and `event_id` in common columns. We can used them to join the 3 dataframes

```{r}
str(test)
```

```{r}
setequal(colnames(test), colnames(train))
```
* test and train have the same variables

```{r}
glimpse(sample_submission)
```

# Filter only installation_id in train_labels

```{r}

train_with_assess <- train %>% 
  filter(installation_id %in% unique(train_labels$installation_id)) 

train_full <- train_with_assess %>%
             left_join(train_labels, by = c("game_session", "installation_id", "title")) 

train_full %>% 
  group_by(installation_id, game_session) %>%
  summarise(s = sum(event_code))
```




```{r}
test <- 
test %>% 
mutate(timestamp = gsub( "T", " ", timestamp)) %>%
mutate(timestamp = gsub( "\\..*Z$", "", timestamp)) %>%
mutate(Date = format(as.POSIXct(timestamp ,format="%Y-%m-%d %H:%M:%S"),"%Y-%m-%d")) %>%
mutate(Months = lubridate::month(ymd(Date), label =TRUE)) %>%
mutate(Time = format(as.POSIXct(timestamp ,format="%Y-%m-%d %H:%M:%S"),"%H:%M")) %>%
mutate(wday = lubridate::wday(ymd(Date), label = TRUE)) %>%
select(Date, wday, Time, everything(), -timestamp)

test %>% head
```





# What is the Goal.

* The goal is to predict the assessement level of each installation_id. The prediction is based of installation_id with history of assessment.

* The file `train_labels.csv` has been provided to show how these groups would be computed on the assessments in the training set. Assessment attempts are captured in `event_code 4100` for all assessments except for `Bird Measurer`, which uses `event_code 4110`. If the attempt was correct, it contains `"correct":true`.

* All installation_id without assessment is not usefull for training.

* `type` column describes if the installation_id are an assessment or other type of App.

```{r}
train_with_assess %>%
  select(type) %>%
  table()
```


```{r}
# # Select installation_id with Assessment from train data 
# train_with_assess <- readRDS("datasets/train_with_assess.rds")
# 
# train_with_assess <- train %>% 
#   filter(type == 'Assessment') %>% 
#   distinct(installation_id) %>% 
#   left_join(train, by = "installation_id")
# 
# saveRDS(train_with_assess, file = "datasets/train_with_assess.rds")
# #fwrite(train_with_assess, file = "datasets/train_with_assess.csv")
```

# Where is the assessment result (correct and incorrect)?



```{r}
train_with_assess <- 
train_with_assess %>% filter(event_code == 4100 | event_code == 4110) %>%
  mutate(event_data = gsub('"', '', event_data)) %>%
  mutate(Correct = str_detect(event_data, pattern = 'correct:true')) %>%
  select(Date,Months, wday, Time, Correct, event_data, everything())

train_with_assess %>% head
```


```{r}
test <- test %>%
  #filter(event_code == 4100 | event_code == 4110) %>% 
  filter(type == 'Assessment') %>%
  mutate(event_data = gsub('"', '', event_data)) %>%
  mutate(Correct = str_detect(event_data, pattern = 'correct:true')) %>%
  select(Date,Months, wday, Time, Correct,event_data, everything()) 

test %>% head 
```




# Accuracy group range 
```{r}
train_labels %>%
  ggplot()+
  aes(x = accuracy_group, y = accuracy)+
  geom_point()
```

* Accuracy_group equal to 0 if accuracy eaqual to 0
* Acuuracy_group equal to 1 if 0 < accuracy < 0.5
* Accuracy_group equal to 2 if 0.5 =< accuracy < 1
* Accuracy_goup equal to 3 if accuracy == 1


# Assessment of tain and Test 
this code try to build the same datafarme like train_labels for Train and Test data

```{r}
train_with_assess %>% 
  group_by(game_session, installation_id, title ) %>%
summarise(num_correct = sum(Correct == TRUE),
          num_incorrect = sum(Correct == FALSE)) %>%
  ungroup %>%
  mutate(accuracy = num_correct/(num_correct+num_incorrect)) %>%
  mutate(accuracy_group = ifelse(accuracy == 0,  0,
          ifelse(accuracy> 0 & accuracy < 0.5,  1,
          ifelse(accuracy >= 0.5 & accuracy < 1,  2,
                  1
          )))) -> train_labels_own

train_labels_own %>% head

```

```{r}
train_labels %>%
  group_by(title)%>%
  summarise(accuracy_group = median(accuracy_group, na.rm = T)) 
```


```{r}
test %>% 
  group_by(game_session, installation_id, title ) %>%
summarise(num_correct = sum(Correct == TRUE),
          num_incorrect = sum(Correct == FALSE)) %>%
    ungroup %>%
  mutate(accuracy = num_correct/(num_correct+num_incorrect)) %>%
  mutate(accuracy_group = ifelse(accuracy == 0,  0,
          ifelse(accuracy> 0 & accuracy < 0.5,  1,
          ifelse(accuracy >= 0.5 & accuracy < 1,  2,
                  1
          )))) -> test_labels

head(test_labels)
```

# get accuracy group by installation_id
```{r}
test %>% 
  group_by( installation_id) %>%
summarise(num_correct = sum(Correct == TRUE),
          num_incorrect = sum(Correct == FALSE)) %>%
    ungroup %>%
  mutate(accuracy = num_correct/(num_correct+num_incorrect)) %>%
  mutate(accuracy_group = ifelse(accuracy == 0,  0,
          ifelse(accuracy> 0 & accuracy < 0.5,  1,
          ifelse(accuracy >= 0.5 & accuracy < 1,  2,
                  1
          )))) -> submission_test


submission_test
```


**Prepare the test by selecting last row of each installation_id ,game_session as we have only 1000 rows in sample_submission.**
```{r}

test %>%
  group_by(Date, installation_id, game_session) %>%
  summarise(n_session = n()) %>%
  arrange(installation_id)
```


```{r}


test %>%
  filter(str_detect(event_data , pattern = "event_code:4100|event_code:4110")) %>% select(installation_id) %>% n_distinct() 

test %>%
  filter(event_code == 4100 | event_code == 4110)  %>% select(installation_id) %>% n_distinct() 

test %>%
  filter(type == 'Assessment')  %>% select(installation_id) %>% n_distinct()
```


```{r}
sample_submission %>% 
  left_join(submission_test, by = "installation_id") 
```


## Explore installation_id
in the descrition we not:

* Each installation corresponds to one child, but in some case there is shared device. It means that 2 children can use the same installation_id.

* The training set contains many installation_ids which never took assessments, whereas every installation_id in the test set made an attempt on at least one assessment.


```{r}
paste("train_with_assess data has: ", train_with_assess$installation_id %>% n_distinct(), "distinct installation_id")
paste("train_labels data has: ", train_labels$installation_id %>% n_distinct(), "distinct installation_id")
paste("test data has: ", test$installation_id %>% n_distinct(), " distinct installation_id")
paste("sample_submission file has : ", sample_submission$installation_id %>% n_distinct(), " distinct installation_id")
```

## Check if test and sample_submission have the same installation_id

```{r}
paste0("Matched installation_id - train_with_assess versus train_labels: ",length(which(unique(train$installation_id) %in% unique(train_labels$installation_id))))
paste0("Matched installation_id - test versus sample_submission: ",length(which(unique(test$installation_id) %in% unique(sample_submission$installation_id))))
paste0("Matched installation_id - test versus train: ",length(which(unique(test$installation_id) %in% unique(train_with_assess$installation_id))))

```

* train and test do not have common installation_id.

* all train_labels installation_id are in train

## Look for which column has common variables

```{r}
paste0("Common event_id between test & train_with_assess: ",length(which(unique(test$event_id) %in% unique(train_with_assess$event_id))))
paste0("Common game_session between test & train_with_assess: ",length(which(unique(test$game_session) %in% unique(train_with_assess$game_session))))
paste0("Common game_time between test & train_with_assess: ",length(which(unique(test$game_time) %in% unique(train_with_assess$game_time))))
paste0("Commen timestamp between test & train_with_assess: ",length(which(unique(test$timestamp) %in% unique(train_with_assess$timestamp))))

```

# Explore Events

```{r}
paste("train_with_assess data has: ", train_with_assess$event_id %>% n_distinct(), "distinct event_id")
#paste("train_labels data has: ", train_labels$event_id %>% n_distinct(), "distinct event_id")
paste("test data has: ", test$event_id %>% n_distinct(), " distinct event_id")
#paste("sample_submission file has : ", sample_submission$event_id %>% n_distinct(), " distinct event_id")

```


```{r}
train_with_assess %>%
  select(timestamp, contains("event")) %>% head
```

```{r}
test %>%
  select(timestamp, contains("event")) %>% head
```

```{r}
specs %>% head
```

## Extract information from JSON columns

the colmuns `event_data` in Train and Test and the column `args` in specs dataframe contain informations in JSON format.

```{r}
#train_event_data <- 
train_with_assess$event_data %>% head(10) %>%
lapply(function(x) fromJSON(gsub('""', "\"", x))) %>%
rbindlist( fill =TRUE)
```
Some case have unregular JSON format which make error message and stop transfomation. 

```{r}
#test_event_data <- 
test$event_data %>% head(10) %>%
lapply(function(x) fromJSON(gsub('""', "\"", x))) %>%
rbindlist( fill =TRUE)

```

```{r}
## Extract value of specific argument
 
```




```{r}
specs$args[3] %>%
lapply(function(x) fromJSON(gsub('""', "\"", x))) %>%
rbindlist( fill =TRUE)

 specs$args %>%
  lapply(function(x) fromJSON(gsub('""', "\"", x))) %>%
  rbindlist(fill =TRUE) %>% head(10)
```


