---
title: "The smile is related or not?"
author: "Karim Mezhoud"
date: '`r Sys.Date()`'
output:
  html_document:
    number_sections: true
    fig_caption: true
    toc: true
    fig_width: 7
    fig_height: 6
    theme: cosmo
    highlight: tango
    code_folding: show #hide
    self_contained: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, error=FALSE, warning = FALSE)
knitr::opts_chunk$set(fig.width=6, fig.height=4) 
```


```{r , message=FALSE}
library(data.table)
library(dplyr)
library(plyr)
library(stringr)
library(EBImage)
library(keras)
library(imager)
```



```{r}
sample_submission <- fread("sample_submission.csv")
head(sample_submission)
```

```{r}
relationship <- data.table::fread("train_relationships.csv",stringsAsFactors = TRUE)
relationship %>%
  filter(p1 == "F0058/MID2")
```
# Processing relationships
The goal is to group and list the folders with kinship.

## Collapse from p1 to p2
```{r}
collaps_P1P2 <- plyr::ddply(relationship,c("p1"),
                            function(df1)paste(df1$p2,
                                               collapse = ","))
colnames(collaps_P1P2) <- c("p1", "p2")
collaps_P1P2 %>%
  head(20) 

```

## Cancatenate subfamilies 1
```{r}
subfamilies1 <- paste(collaps_P1P2$p1, collaps_P1P2$p2, sep=",")
## extract subfamilies with more than two persons
subfamilies1_  <- subfamilies1[stringr::str_detect(subfamilies1, "\\w\\d*\\/\\w*\\d*,\\w\\d*\\/\\w*\\d*")]
subfamilies1_ %>%
  head(20)
```

## Collapse from p2 to p1
```{r}
collaps_P2P1 <- plyr::ddply(relationship,c("p2"),
                            function(df1)paste(df1$p1,
                                               collapse = ","))
colnames(collaps_P2P1) <- c("p2", "p1")
collaps_P2P1 %>%
  head(20)
```

## Cancatenate subfamilies 2

```{r}
subfamilies2 <- paste(collaps_P2P1$p1, collaps_P2P1$p2, sep=",")
## extract subfamilies with more than two persons
subfamilies2_  <- subfamilies2[stringr::str_detect(subfamilies2, "\\w\\d*\\/\\w*\\d*,\\w\\d*\\/\\w*\\d*")]
subfamilies2_ %>%
  head(20)
```


```{r}
families <- unique(c(subfamilies1_, subfamilies2_))
length(families)

```

## Index of person in the same family
```{r}
family_idx <- 
  unique(unlist(strsplit(families, ","))) %>% 
  str_extract("\\w\\d*") %>% 
  as.factor() %>%
  as.numeric()

family_idx
```

## Collapse Persons by family index
```{r}
fam <- as.data.frame(cbind(Person = unique(unlist(strsplit(families, ","))),family_idx))
fam %>%
  head(20)
```


```{r}
unique_families <- plyr::ddply(fam,c("family_idx"),
                               function(df1)paste(df1$Person,
                                                  collapse = ","))
colnames(unique_families) <- c("index", "Persons with kinship relationships")
DT::datatable(unique_families) %>%
  DT::formatStyle( colnames(unique_families), color = 'black')
```

**There are 470 kinship relationships **


```{r}
unique_families$`Persons with kinship relationships`[1:10]

#length(str_split(unique_families$`Persons with kinship relationships`[10],pattern = ","))
```

# Generate all combinations `is_related` and `no_related` from each family
```{r, warning=FALSE}
person <- unique_families$`Persons with kinship relationships`[1] %>%
  str_split(pattern = ",", simplify = TRUE)

person_permut <-
  gtools::permutations(n = length(person),
                       r = 2,
                       v = person,
                       repeats.allowed = FALSE) 

## split the list of related person by comma
split_person <-    lapply(unique_families[,2], function(x) str_split(x, pattern = ",", simplify = TRUE))
## get all possible permutation between persons for each family
split_person_permut <- lapply(split_person, function(x)  gtools::permutations(n = length(x),
                                                                              r = 2,
                                                                              v = x,
                                                                              repeats.allowed = FALSE))
## merge the list of dataframe to dataframe
all_person_permut <- plyr::ldply(split_person_permut, data.frame)
colnames(all_person_permut) <- c("p1", "p2")

## 
no_related <- dplyr::anti_join( all_person_permut,relationship, by = c("p1", "p2"))
## 
is_related <- dplyr::semi_join( all_person_permut,relationship, by = c("p1", "p2"))

## clean related person but in opposite format: p1 -> p2 or p2 -> p1
## we need to rename zero (p2, p1) and omit commun row with one.
colnames(no_related) <- c("p2", "p1")

## There is two way of relation p1 -> p2 and p2 -> p1

## extract the second way of relation from no_related df
is_related_inverse <- dplyr::semi_join(no_related, is_related, by= c("p1", "p2"))
## remove related person p2 -> p1
no_related_inverse <- dplyr::anti_join(no_related, is_related, by = c("p1", "p2"))

colnames(is_related_inverse) <- c("p1", "p2")
colnames(no_related_inverse) <- c("p1", "p2")

## Merge the 2 ways of relation
is_related_all <- rbind(is_related, is_related_inverse)


no_related_all <-  no_related_inverse

## Annotate 0, 1
is_related_all <- cbind(is_related_all, is_related = 1)
no_related_all <- cbind(no_related_all, is_related = 0)

head(is_related_all)
```
# Check folder and files availabilities
The goal is to clean up `is_related_all` and `no_related_all` from not existing folders or images.

```{r}
is_related_all <-
is_related_all %>%
  filter(file.exists(paste0("train/", p1))) %>%
  filter(file.exists(paste0("train/", p2))) %>%
  filter(!length(list.files(paste0("train/", p1))) == 0) %>%
  filter(!length(list.files(paste0("train/", p2))) == 0)


no_related_all <-
  no_related_all %>%
  filter(file.exists(paste0("train/", p1))) %>%
  filter(file.exists(paste0("train/", p2))) %>%
  filter(!(length(list.files(paste0("train/", p1))) == 0)) %>%
  filter(!(length(list.files(paste0("train/", p2))) == 0))
  
is_related_all %>%
  filter(p1 == "F0460/MID8") 

length(list.files("train/F0460/MID8/"))
```

**Note** `filter` can not detect particular empty folders

```{r}
kinships_all <- 
  is_related_all %>%
  bind_rows(no_related_all) %>%
  arrange(by = p1)

kinships_all %>%
  head(8)
```


# Images Data processing 

# Read and reshape related images and label from dataframe

```{r}
# #inputs are 2 images and one category
# # df shape : 1 row ,3 columns
# loadinputs <- function(df){
#   
#   pathfile1 <-  paste0("train/", df[[1]])
#   pathfile2 <-  paste0("train/", df[[2]])
#   #label     <-  keras::to_categorical(df[[3]])
#   label     <-           as.factor(df[[3]])
#   
#   # sample one image
#   name1 <- sample(x = list.files(pathfile1), size = 1, replace = TRUE)
#   name2 <- sample(x = list.files(pathfile2), size = 1, replace = TRUE)
#   
#   # load image 
#   img1 <- EBImage::readImage(paste0(pathfile1,"/",name1))
#   img2 <- EBImage::readImage(paste0(pathfile2,"/",name2))
#   # resize images
#   img1 <-   EBImage::resize(img1, 28, 28)
#   img2 <-   EBImage::resize(img2, 28, 28)
#   ## reshape images
#   img1 <- keras::array_reshape(img1, c(28,28,3))
#   img2 <- keras::array_reshape(img2, c(28,28,3))
#   
#   train1 <- rbind(train1, img1)
#   train2 <- rbind(train2, img2)
#   trainy <- label
#   ls <- list(train1, train2, label)
#   #train1 <- t(train1)
#   return(train1)
#   
#   ## return as a list
#   #ls <- list(img1,img2,label)
#   #names(ls) <- c(basename(pathfile1), basename(pathfile2), "label")
#   #return(ls)
# }
# 
# train1 <- NULL
# train2 <- NULL
# trainy <- NULL
# lklk <- apply(is_related_all[1:38,],1, function(row) loadinputs(row))
# 
# ## family label
# #names(lklk) <- apply(is_related_all[1:3,], 1, function(x) 
# #                    names(x) <- stringr::str_split(string = x[[1]],
# #                                                  pattern = '/',
# 
# ## Unname dimnames attributes of the matrix. We can not do this in apply function.
# lklk <- unname(lklk)
# lklk <- t(lklk)
# str(lklk)
```

## get Train matrix from images
In our case we need as input two images and one label. The setting of the model will accept 2 matrices as `trainx` and a label as `trainy`.

```{r}
get_train_matrix <- function(df, col){
   pathfolder <-  paste0("train/", df[col])

    ## Check if the path exists && the folder is not empty
  if(file.exists(pathfolder) && !length(list.files(pathfolder)) == 0){

  set.seed(123)
  # sample one image
   name1 <- sample(x = list.files(pathfolder), size = 1, replace = TRUE)

  # load image
  img1 <- imager::load.image(paste0(pathfolder,"/",name1))
  #img1 <- EBImage::readImage(paste0(pathfolder,"/",name1))

  # resize images
  img1 <- imager::resize(img1, 32, 32)
  #img1 <-   EBImage::resize(img1, 32, 32)

  ## reshape images
  img1 <- keras::array_reshape(img1, c(32,32,3))

  trainx <- rbind(trainx, img1)

  }else{
    to_delete <<- c(to_delete, paste0(pathfolder))
  }
   return(trainx)
}
to_delete <- NULL
trainx <- NULL
star <- Sys.time()
trainx1 <-  apply( kinships_all, 1, function(row) get_train_matrix(row, 1))
end <- Sys.time()
print(end - star)
# Time difference of 2.872232 mins

#trainx1 <- NULL
#star <- Sys.time()
## works but same think can not remove attributes in each matrix
#trainx1 <-purrr::map(kinships_all, .f = function(row) get_train_matrix(row, 1))
#end <- Sys.time()
#print(end - star)

## clean empty list of not existing folder or images
#trainx <- rlist::list.clean(trainx, recursive = TRUE)
## works also
#trainx <-  trainx[unlist(lapply(trainx, length) != 0)]
 

trainx1 <- ldply(trainx1, function(x) x <- unname(x))
trainx1 <- t(trainx1)
#trainx1 <-lapply(trainx1, function(x) unname(x)) 
#str(lapply(trainx1, function(x) unname(x)) )
trainx1 <- unname(trainx1)
str(trainx1)

```

```{r}
# store empty folders that not filtred 
to_delete1 <- to_delete %>% unique() %>%
  stringr::str_split(pattern ="/", simplify = TRUE, n = 2)
  
  
to_delete1 <- to_delete1[,2]
to_delete1
```

## Loop is not the best way but I am sure that the str of images is conserved.
```{r}

get_trainx <- function(df){
  
  pb = txtProgressBar(min = 0, max = length(df), initial = 0)
  
  trainx <- NULL
  
  for(i in seq_len(length(df))){
    
      setTxtProgressBar(pb,i,
                       title=paste(round(i/length(df) *100, 0),"% done"))

    pathfolder <-  paste0("train/", df[i])
    
    ## Check if the path exists && the folder is not empty
    if(file.exists(pathfolder) && !length(list.files(pathfolder)) == 0){
      
      set.seed(123)
      # sample one image
      name1 <- sample(x = list.files(pathfolder), size = 1, replace = TRUE)
      
      # load image 
      #img1 <- imager::load.image(paste0(pathfolder,"/",name1))
      img1 <- EBImage::readImage(paste0(pathfolder,"/",name1))
      
      # resize images
      #img1 <- imager::resize(img1, 32, 32)
      img1 <-   EBImage::resize(img1, 32, 32)
      
      ## reshape images
      img1 <- keras::array_reshape(img1, c(32,32,3))
      
      trainx <- rbind( trainx, img1)
      
    }
  }
  return(trainx)
}

star <- Sys.time()
trainx <- get_trainx(kinships_all[1:10,1])
end <- Sys.time()
print(end - star)
# difference of 27.67416 mins
trainx <- unname(trainx)
str(trainx)
```
Just to verify if the matrix structure is the same afetr looping and applying. 

## get trainx 2
```{r}
to_delete <- NULL
trainx <- NULL
star <- Sys.time()
trainx2 <-  apply( kinships_all, 1, function(row) get_train_matrix(row, 2))
end <- Sys.time()
print(end - star)


trainx2 <- ldply(trainx2, function(x) x <- unname(x))
trainx2 <- t(trainx2)
trainx2 <- unname(trainx2)
str(trainx2)
```

```{r}
## store to_delete2
to_delete2 <- to_delete %>% unique() %>%
  stringr::str_split(pattern ="/", simplify = TRUE, n = 2)

to_delete2 <- to_delete2[,2]
to_delete2
```


## re--filter kinships_all

```{r}
kinships_all_filtered <-
  kinships_all %>%
  filter(p1 != to_delete1[1]) %>%
  filter(p1 != to_delete1[2]) %>%
  filter(p2 != to_delete2[1]) %>%
  filter(p2 != to_delete2[2])

kinships_all_filtered %>%
  filter(p1== to_delete[1])
dim(kinships_all_filtered)
```

**Note**
We found 12.474 rows while trainx1 and trainx2 have 12.497.
There are also 2 empty folders that we have to identify.
We will re-run `get_train_matrix` with the new filtered a list of folders.

```{r}
to_delete <- NULL
trainx <- NULL
star <- Sys.time()
trainx2 <-  apply( kinships_all_filtered, 1, function(row) get_train_matrix(row, 2))
end <- Sys.time()
print(end - star)

trainx2 <- t(trainx2)
# trainx2 <- as.matrix(ldply(trainx2, function(x) x <- unname(x)))
# trainx2 <- unname(trainx2)
 str(trainx2)


## store to_delete2
to_delete2 <- to_delete %>% unique() %>%
  stringr::str_split(pattern ="/", simplify = TRUE, n = 2)

to_delete2 <- to_delete2[,2]
to_delete2
```

```{r}
to_delete <- NULL
trainx <- NULL
star <- Sys.time()
trainx1 <-  apply( kinships_all_filtered, 1, function(row) get_train_matrix(row, 1))
end <- Sys.time()
print(end - star)

 trainx1 <- t(trainx1)
# trainx1 <- as.matrix(ldply(trainx1, function(x) x <- unname(x)))
# trainx1 <- unname(trainx1)
str(trainx1)


## store to_delete2
to_delete1 <- to_delete %>% unique() %>%
  stringr::str_split(pattern ="/", simplify = TRUE, n = 2)

to_delete1 <- to_delete1[,2]
to_delete1
```


##  get Labels
```{r}
Labels <- keras::to_categorical(kinships_all_filtered[,3])
str(Labels)
```


# Keras tentative
## sample test and Train data
```{r}
set.seed(982)
idx <- seq_len(nrow(trainx1))
idx_train <- sample(idx, max(idx) * 0.8, replace = FALSE)
  
  train_x1 <- trainx1[idx_train,]
  test_x1 <- trainx1[-idx_train,]
  train_x2 <- trainx2[idx_train,]
  test_x2 <- trainx2[-idx_train,]
  
  Labels_train <- keras::to_categorical(Labels[idx_train])
  Labels_test <- keras::to_categorical(Labels[idx[-idx_train]])
 
  dim(train_x1)
str(Labels_test)
```
## Set the model
```{r}

# input1 <- layer_input(shape =  c(28, 28, 3))
# input2 <- layer_input(shape = c(28, 28, 3))
# 
# conv1 <- input1 %>% 
#   layer_conv_2d(filters = 7, kernel_size = c(1, 1), padding='same', activation='relu') %>% 
#   layer_conv_2d(filters = 7, kernel_size = c(3, 3), padding='same', activation='relu')
# 
# conv2 <- input2 %>% 
#   layer_conv_2d(filters = 7, kernel_size = c(1, 1), padding='same', activation='relu') %>% 
#   layer_conv_2d(filters = 7, kernel_size = c(3, 3), padding='same', activation='relu')
# 
# 
# output <- 
#   layer_concatenate(c(conv1, conv2), axis = 1) %>%
#   layer_dense(units = 256, activation = 'relu', input_shape = c(3072)) %>% # 28 x 28 x 3
#   layer_dense(units = 32, activation = 'relu') %>%
#   layer_dense(units = 2, activation = 'softmax') # we have two output 0,1


input1 <- layer_input(shape =  c(3072))
input2 <- layer_input(shape =  c(3072))

#conv1 <- layer_conv_2d( filters = 4, input_shape = c(28, 28, 3), kernel_size = 3)
#conv2 <- layer_conv_2d( filters = 4, input_shape = c(28, 28, 3), kernel_size = 3)

output <-
  layer_concatenate(list(input1, input2)) %>% 
 #  layer_dense(units = 512, activation = 'relu',
  #           kernel_regularizer = regularizer_l2(l = 0.001)) %>%
  # layer_dropout(rate = 0.1) %>% 
  #    layer_dense(units = 256, activation = 'relu',
   #              kernel_regularizer = regularizer_l2(l = 0.001)) %>%
  #  layer_dropout(rate = 0.05) %>%
       layer_dense(units = 100, activation = 'relu',
               kernel_regularizer = regularizer_l2(l = 0.001)) %>%
  #  layer_dropout(rate = 0.02) %>%
      layer_dense(units = 50, activation = 'relu',
             kernel_regularizer = regularizer_l2(l = 0.001)) %>%
 #layer_dropout(rate = 0.01) %>%

  layer_dense(units = 2, activation = 'sigmoid') # we have two output 0,1
  #  layer_flatten() %>%
  #layer_dense(units = 2, activation = "softmax")


model <-
  keras_model(inputs = list(input1, input2), outputs = output)

model %>% compile(optimizer =  "adam", # , optimizer_rmsprop()
                  loss =  'binary_crossentropy',
                  metrics = 'accuracy')

summary(model)

```

## train the model
```{r}
history <- 
  model %>% 
  fit(list(train_x1, train_x2),
      Labels_train, 
      epochs = 100,
      batch_size = 16,
      verbose = 1,             # 2
      validation_split = 0.2) # list(test_x1, Labels_test),

plot(history)



```



## fit_generator

```{r}
sampling_generator <- function(train_x1, train_x2, Labels_train, batch_size){
    function() {
    rows <- sample(1:nrow(train_x1), batch_size, replace = TRUE)
    
  ls_train <-  list(train_x1[rows,],
         train_x2[rows,],
         Labels_train[rows,])
    #return(ls_train)
    }

}

model %>% 
  keras::fit_generator(sampling_generator(train_x1, 
                                          train_x2,
                                          Labels_train,
                                          batch_size = 128),
            steps_per_epoch = nrow(train_x1) / 128, epochs = 10)

```

##  Evaluate
```{r}
score <- model %>% evaluate(
  list(test_x1, test_x2),
  Labels_test,
  verbose = 0
)

# Output metrics
cat('Test accuracy:', score[[2]], '\n')
cat('Test loss:', score[[1]], '\n')

```


# Prediction
```{r}
sample_submission <- fread("sample_submission.csv")
head(sample_submission)
```

## get testx1 and testx2 from sample_submission file
```{r}

sample_submission[, c("p1", "p2") := tstrsplit(img_pair, "-", fixed=TRUE)]
sample_submission %>%
  head()
```




```{r}
get_test_matrix <- function(df, col){
   pathfile <-  paste0("test/", df[col])

    ## Check if the path exists && the folder is not empty
  if(file.exists(pathfile)){

  # load image
  img <- EBImage::readImage(pathfile)

  # resize images
  img <-   EBImage::resize(img, 32, 32)

  ## reshape images
  img <- keras::array_reshape(img, c(32,32,3))

  testx <- rbind(testx, img)

  }else{
    to_delete <<- c(to_delete, paste0(pathfile))
  }
   return(testx)
}

to_delete <- NULL
testx <- NULL
star <- Sys.time()
testx1 <-  apply(sample_submission, 1, function(row) get_test_matrix(row, 3))
end <- Sys.time()
print(end - star)
testx1 <- t(testx1)
str(testx1)
to_delete
```

```{r}
to_delete <- NULL
testx <- NULL
star <- Sys.time()
testx2 <-  apply(sample_submission, 1, function(row) get_test_matrix(row, 4))
end <- Sys.time()
print(end - star)
testx2 <- t(testx2)
str(testx2)
to_delete
```


```{r}
is_related <- predict(model, list(testx1, testx2))
is_related %>%
  head()
```

# submission
```{r}
dt_submission <- data.table(
  img_pair = sample_submission$img_pair,
  is_related = round(is_related[,1], digits = 1))
head(dt_submission)
```

```{r}
fwrite(dt_submission, file = "submission.csv")
```


# save model
```{r}
#model %>% save_model_hdf5("my_model.h5")
#new_model <- load_model_hdf5("my_model.h5")
```

