---
title: "Santandar Costumer Transaction Prediction with xgboost"
author: "Karim Mezhoud"
date: '`r Sys.Date()`'
output:
  html_document:
    number_sections: true
    fig_caption: true
    toc: true
    fig_width: 7
    fig_height: 6
    theme: cosmo
    highlight: tango
    code_folding: show #hide
    self_contained: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

# Load packages
```{r comment=FALSE, warning=FALSE}
library(data.table)
library(dplyr)
library(rsample)
library(xgboost)
require(Ckmeans.1d.dp)
library(ggplot2)

```

# Load Data

```{r}
train <- fread(file = "train.csv")
train[1:10,1:14]
```

# Split Into Train/Test Sets
```{r}
set.seed(100)
train_test_split <- rsample::initial_split(train, prop = 0.8)
train_test_split
```

We can retrieve our training and testing sets using training() and testing() functions.

```{r}
# Retrieve train and test sets
train_8 <- rsample::training(train_test_split)
test_2  <- rsample::testing(train_test_split)
train_8[1:10, 1:14]
```


# Optimize features

Here, we can see after how many rounds, we achieved the smallest test error:

```{r}
dtrain <- xgb.DMatrix(as.matrix(train_8[, -c(1,2)]), 
                      label = as.numeric(train_8$target))
dtest <- xgb.DMatrix(as.matrix(test_2[, -c(1,2)]), 
                      label = as.numeric(test_2$target))

params <- list(max_depth = 2, 
               objective = "binary:logistic",
               silent = 0)

watchlist <- list(train = dtrain, eval = dtest)




cv_model <- xgb.cv(params = params,
                   data = dtrain,
                   nrounds = 100,
                   eta = 1,
                   watchlist = watchlist,
                   nfold = 5,
                   verbose = TRUE,
                   prediction = TRUE) # prediction of cv folds


# eval_model <- xgb.train(data=dtrain, 
#                         max.depth=2,
#                         eta = 0.5,
#                         nthread = 4, 
#                         nrounds=50,
#                         watchlist=watchlist,
#                         eval.metric = "error", 
#                         eval.metric = "logloss", 
#                         objective = "binary:logistic")
```

# Train the model
## Tree boosting
```{r}
tme <- Sys.time()
xgboost_tree <- xgb.train(data = dtrain, 
                         max_depth = 2, 
                         eta = 0.1, 
                         objective = "binary:logistic",
                         nthread = 2,
                         nrounds = 1500,
                         eval_metric = "auc",
                         verbose = TRUE)
Sys.time() - tme
```

## Linear boosting (algorithme)
```{r}

xgboost_linear <- xgboost(data = as.matrix(train_8[, c(-1,-2)]), 
                         label = as.numeric(train_8$target),
                         booster = "gblinear",
                         learning_rates = 1,
                         max_depth = 2, 
                         objective = "binary:logistic", 
                         nrounds = 100, 
                         verbose = TRUE)



```


## Extract important features from tree xgboost model
```{r}
features <- colnames(train_8[, c(-1,-2)])
importance_matrix_tree <- xgb.importance(features, model = xgboost_tree)
importance_matrix_tree
```

## Extract important features from linear xgboost model
```{r}
features <- colnames(train_8[, c(-1,-2)])
importance_matrix_lm <- xgb.importance(features, model = xgboost_linear)
importance_matrix_lm %>%
arrange(desc(Weight)) %>%
  head(30)
```

## plot important features (linear model)

```{r}
xgb.ggplot.importance(importance_matrix_lm[1:30,]) +
ggplot2::theme_minimal()
```

## plot important features (Tree model)
```{r}
xgb.ggplot.importance(importance_matrix_tree[1:30,]) +
ggplot2::theme_minimal()
```

# Testing models
## Testing linear model
```{r}
pred_lm_2 <- predict(xgboost_linear, as.matrix(test_2[, c(-1,-2)]))
head(pred_lm_2)
```

## testing Tree model
```{r}
pred_tree_2 <- predict(xgboost_tree, as.matrix(test_2[, c(-1,-2)]))
head(pred_tree_2)
```

## Transform propability to binary classification
```{r}
prediction_lm_2 <- as.numeric(pred_lm_2 > 0.6)
table(prediction_lm_2)
```
```{r}
prediction_tree_2 <- as.numeric(pred_tree_2 > 0.6)
table(prediction_tree_2)
```

## Confusion Matrix for linear model
```{r}

data.frame(prediction = as.numeric(prediction_lm_2),
         label = as.numeric(test_2$target)) %>%
  count(prediction, label)
```

## Confusion matric for Tree model
```{r}
data.frame(prediction = as.numeric(prediction_tree_2),
         label = as.numeric(test_2$target)) %>%
  count(prediction, label)
```

# Prediction

## Load test data
```{r}
test <- fread(file = "test.csv")
test[1:10,1:14]
```

## Prediction with linear xgboost model
```{r}

pred_lm <- predict(xgboost_linear, as.matrix(test[,-1]))
head(pred_lm)
```

## Prediction with Tree xgboost model

```{r}
pred_tree <- predict(xgboost_tree, as.matrix(test[,-1]))
head(pred_tree)
```

These numbers doesn’t look like binary classification `{0,1}. We need to perform a simple transformation before being able to use these results.

# Transform the regression in a binary classification
If we think about the meaning of a regression applied to our data, the numbers we get are probabilities that a datum will be classified as 1. Therefore, we will set the rule that if this probability for a specific datum is > 0.5 then the observation is classified as 1 (or 0 otherwise).

```{r}
prediction_lm <- as.numeric(pred_lm > 0.5)
table(prediction_lm)
```

```{r}
prediction_tree <- as.numeric(pred_tree > 0.5)
table(prediction_tree)

```

# submission

We select linear model for submission. Linear model prediction has about 1% rate of `1`, like in the train dataset.

```{r}
dt_submission <- data.frame(
  ID_code = test[,1],
  target = pred_tree
)
head(dt_submission)
```

```{r}
#fwrite(dt_submission[,c(1,2)], "submission.csv")
```


