---
title: "Dealing with unbalanced data in machine learning"
author: "Karim Mezhoud"
date: '`r Sys.Date()`'
output:
  html_document:
    number_sections: true
    fig_caption: true
    toc: true
    fig_width: 14
    fig_height: 8
    theme: cosmo
    highlight: tango
    code_folding: show #hide
    self_contained: no

---

```{r}
library(tidyverse)
library(caret)
library(data.table)
library(randomForest)
```

[shiring github](https://shiring.github.io/machine_learning/2017/04/02/unbalanced)

```{r}
bc_data <- read.table("breast-cancer-wisconsin.data.txt", 
                      header = FALSE, 
                      sep = ",")


colnames(bc_data) <- c("sample_code_number", 
                       "clump_thickness", 
                       "uniformity_of_cell_size", 
                       "uniformity_of_cell_shape", 
                       "marginal_adhesion", 
                       "single_epithelial_cell_size", 
                       "bare_nuclei", 
                       "bland_chromatin", 
                       "normal_nucleoli", 
                       "mitosis", 
                       "classes")

bc_data$classes <- ifelse(bc_data$classes == "2", "benign",
                          ifelse(bc_data$classes == "4", "malignant", NA))


bc_data$classes <- as.factor(bc_data$classes)
bc_data[bc_data == "?"] <- NA

# how many benign and malignant cases are there?
summary(bc_data$classes)
```

## Missing values are imputed with the mice package.

```{r}
# impute missing data
library(mice)

bc_data[,2:10] <- apply(bc_data[, 2:10], 2, function(x) as.numeric(as.character(x)))
dataset_impute <- mice(bc_data[, 2:10],  print = FALSE)
bc_data <- cbind(bc_data[, 11, drop = FALSE], mice::complete(dataset_impute, 1))

bc_data$classes <- as.factor(bc_data$classes)

nrow(bc_data[is.na(bc_data), ])
```


# Modeling the original unbalanced data

```{r}
set.seed(42)
index <- caret::createDataPartition(bc_data$classes, p = 0.7, list = FALSE)

train_data <- bc_data[index, ]
test_data  <- bc_data[-index, ]

set.seed(42)
model_rf <- caret::train(classes ~ .,
                         data = train_data,
                         method = "rf",
                         preProcess = c("scale", "center"),
                         trControl = trainControl(method = "repeatedcv", 
                                                  number = 10, 
                                                  repeats = 10, 
                                                  verboseIter = FALSE)
                         )




final <- data.frame(actual = test_data$classes,
                    predict(model_rf, newdata = test_data, type = "prob"))

final$predict <- as.factor(ifelse(final$benign > 0.5, "benign", "malignant"))

cm_original <- confusionMatrix(final$predict, final$actual)

cm_original
```

# Under-sampling


```{r}


ctrl <- trainControl(method = "repeatedcv", 
                     number = 10, 
                     repeats = 10, 
                     verboseIter = FALSE,
                     sampling = "down")

set.seed(42)
model_rf_under <- caret::train(classes ~ .,
                         data = train_data,
                         method = "rf",
                         preProcess = c("scale", "center"),
                         trControl = ctrl)

final_under <- data.frame(actual = test_data$classes,
                    predict(model_rf_under, newdata = test_data, type = "prob"))

final_under$predict <- as.factor(ifelse(final_under$benign > 0.5, "benign", "malignant"))

cm_under <- confusionMatrix(final_under$predict, final_under$actual)

cm_under
```


# Oversampling

```{r}

ctrl <- trainControl(method = "repeatedcv", 
                     number = 10, 
                     repeats = 10, 
                     verboseIter = FALSE,
                     sampling = "up")

set.seed(42)
model_rf_over <- caret::train(classes ~ .,
                         data = train_data,
                         method = "rf",
                         preProcess = c("scale", "center"),
                         trControl = ctrl)

final_over <- data.frame(actual = test_data$classes,
                          predict(model_rf_over, newdata = test_data, type = "prob"))

final_over$predict <- as.factor(ifelse(final_over$benign > 0.5, "benign", "malignant"))

cm_over <- confusionMatrix(final_over$predict, final_over$actual)

cm_over

```


# ROSE
Besides over- and under-sampling, there are hybrid methods that combine under-sampling with the generation of additional data. Two of the most popular are ROSE and SMOTE.

```{r}
library(ROSE)

ctrl <- trainControl(method = "repeatedcv", 
                     number = 10, 
                     repeats = 10, 
                     verboseIter = FALSE,
                     sampling = "rose")

set.seed(42)
model_rf_rose <- caret::train(classes ~ .,
                              data = train_data,
                              method = "rf",
                              preProcess = c("scale", "center"),
                              trControl = ctrl)

final_rose <- data.frame(actual = test_data$classes,
                         predict(model_rf_rose, newdata = test_data, type = "prob"))

final_rose$predict <- as.factor(ifelse(final_rose$benign > 0.5, "benign", "malignant"))

cm_rose <- confusionMatrix(final_rose$predict, final_rose$actual)

cm_rose
```


# SMOTE

```{r}

ctrl <- trainControl(method = "repeatedcv", 
                     number = 10, 
                     repeats = 10, 
                     verboseIter = FALSE,
                     sampling = "smote")

set.seed(42)
model_rf_smote <- caret::train(classes ~ .,
                              data = train_data,
                              method = "rf",
                              preProcess = c("scale", "center"),
                              trControl = ctrl)

final_smote <- data.frame(actual = test_data$classes,
                         predict(model_rf_smote, newdata = test_data, type = "prob"))

final_smote$predict <- as.factor(ifelse(final_smote$benign > 0.5, "benign", "malignant"))

cm_smote <- confusionMatrix(final_smote$predict, final_smote$actual)
```




## over under sampling by UBL

Problem with memory!

```{r}
#library(UBL)
#train_ubl <- UBL::AdasynClassif(target~., test_set, beta = 1)

#sum(is.na(train))

## Does not work
  
```


## over under sampling by DMwR

```{r fig.height=6, fig.width=7 }
train <- fread("../cern_particle/train.csv")

train$target <- as.factor(train$targetarget)

train_dmwr_1 <- DMwR::SMOTE(target~., train) #, perc.over = 600,perc.under=100

train_dmwr_2 <- DMwR::SMOTE(target~., train, perc.over = c(500),perc.under=100)

p1 <- train %>%
  mutate(target = as.factor(target)) %>%
  group_by(target) %>%
  summarise(Count = n()) %>%
  ggplot()+
  aes(x = target, y = Count, fill= target) +
  geom_col()+
  geom_text(aes(label = percent(Count/sum(Count))), vjust = -0.5)+
  geom_text(aes(label = Count), vjust = +1)  

p2 <- train_dmwr_1 %>%
  mutate(target = as.factor(target)) %>%
  group_by(target) %>%
  summarise(Count = n()) %>%
  ggplot()+
  aes(x = target, y = Count, fill= target) +
  geom_col()+
  geom_text(aes(label = percent(Count/sum(Count))), vjust = -0.5)+
  geom_text(aes(label = Count), vjust = +1)

p3 <- train_dmwr_2 %>%
  mutate(target = as.factor(target)) %>%
  group_by(target) %>%
  summarise(Count = n()) %>%
  ggplot()+
  aes(x = target, y = Count, fill= target) +
  geom_col()+
  geom_text(aes(label = percent(Count/sum(Count))), vjust = -0.5)+
  geom_text(aes(label = Count), vjust = +1)

gridExtra::grid.arrange(p1, p2, p3, nrow = 3)
```


## Tunning over under sampling by DMwR

```{r}
train_muon <- train %>% 
  mutate(target = as.character(target)) %>%
  filter(target != "muon") %>% 
  mutate(target = as.factor(target))

train_pion <-  train %>% 
  mutate(target = as.character(target)) %>%
  filter(target != "pion") %>% 
  mutate(target = as.factor(target))

train_dmwr_muon <- DMwR::SMOTE(target~., train_muon)

train_dmwr_pion <- DMwR::SMOTE(target~., train_pion, perc.over = 5000 ,perc.under = 5000)

train_dmwr_pion_2 <- DMwR::SMOTE(target~., train_dmwr_pion)



train_dmwr_pion_2 %>%
  mutate(target = as.factor(target)) %>%
  group_by(target) %>%
  summarise(Count = n()) %>%
  ggplot()+
  aes(x = target, y = Count, fill= target) +
  geom_col()+
  geom_text(aes(label = percent(Count/sum(Count))), vjust = -0.5)+
  geom_text(aes(label = Count), vjust = +1)

```



## Manual under-sampling
use size number of smaller class.

```{r}
classes <- unique(train$target)

fin <- NULL
min_samp <- table(train$target)[which.min(table(train$target))]

for (i in classes) {

sub <- subset(train, target == i)

sam <- sub[sample(nrow(sub), min_samp), ]

fin <- rbind(fin, sam)
}

fin %>%
  mutate(target = as.factor(target)) %>%
  group_by(target) %>%
  summarise(Count = n()) %>%
  ggplot()+
  aes(x = target, y = Count, fill= target) +
  geom_col()+
  geom_text(aes(label = percent(Count/sum(Count))), vjust = -0.5)+
  geom_text(aes(label = Count), vjust = +1)

table(fin$target)
```

## Manual sampling using dplyr 


```{r}
table(train$target)

train %>%
  group_by(target) %>%
  sample_n(100000, replace = TRUE) %>%
  ungroup %>%
  mutate(target = as.factor(target)) %>%
  group_by(target) %>%
  summarise(Count = n()) %>%
  ggplot()+
  aes(x = target, y = Count, fill= target) +
  geom_col()+
  geom_text(aes(label = percent(Count/sum(Count))), vjust = -0.5)+
  geom_text(aes(label = Count), vjust = +1)
```

# Predictions

Now let’s compare the predictions of all these models:

```{r}
models <- list(original = model_rf,
                       under = model_rf_under,
                       over = model_rf_over,
                       smote = model_rf_smote,
                       rose = model_rf_rose)

resampling <- resamples(models)
bwplot(resampling)

```


# Models comparaison

```{r}
library(dplyr)
comparison <- data.frame(model = names(models),
                         Sensitivity = rep(NA, length(models)),
                         Specificity = rep(NA, length(models)),
                         Precision = rep(NA, length(models)),
                         Recall = rep(NA, length(models)),
                         F1 = rep(NA, length(models)))

for (name in names(models)) {
  label <- get(paste0("cm_", name))
  
  comparison[comparison$model == name, ] <- comparison %>%
    filter(model == name) %>%
    mutate(Sensitivity = label$byClass[["Sensitivity"]],
           Specificity = label$byClass[["Specificity"]],
           Precision = label$byClass[["Precision"]],
           Recall = label$byClass[["Recall"]],
           F1 = label$byClass[["F1"]])
}

library(tidyr)
comparison %>%
  gather(x, y, Sensitivity:F1) %>%
  ggplot(aes(x = x, y = y, color = model)) +
    geom_jitter(width = 0.2, alpha = 0.5, size = 3)
```

# Function to pull all code and return plot to compare models

```{r}


get_cms <- function(ctrl_method = "repeatedcv",
                                 ctrl_sampling,
                                 target,
                                 data,
                                 train_method = "rf",
                                 preProcess = c("scale", "center"),
                                 test
                           ){
  
  ctrl <- trainControl(method = ctrl_method, 
                       number = 10, 
                       repeats = 10, 
                       verboseIter = FALSE,
                       sampling = ctrl_sampling)
  
  model <- caret::train(as.formula(paste0(target,"~ .", sep = " ")),
                        data = data,
                        method = train_method,
                        preProcess = preProcess,
                        trControl = ctrl)
  
final <- data.frame(actual = test[[target]],
                         predict(model, newdata = test, type = "prob"))

final$predict <- as.factor(ifelse(final$benign > 0.5, "benign", "malignant"))

cm <- confusionMatrix(final$predict, final$actual)
  
  return(cm)
  
}


sampling_list <- c( "down", "up") # "none", , "rose", "smote"

cms <- lapply(sampling_list, function(x) x <- get_cms(ctrl_sampling = x,
                                            target = "classes",
                                            data = train_data,
                                            test = test_data))

plot_cms <- function(ctrl_method = "repeatedcv",
                                 ctrl_sampling,
                                 target,
                                 train,
                                 train_method = "rf",
                                 preProcess = c("scale", "center"),
                                 test){



cms <- lapply(sampling_list, function(x) x <- get_cms(ctrl_sampling = x,
                                            target = target,
                                            data = train,
                                            test = test))

names(cms) <- sampling_list
  
comparison <- data.frame(model = names(cms),
                         Sensitivity = rep(NA, length(cms)),
                         Specificity = rep(NA, length(cms)),
                         Precision = rep(NA, length(cms)),
                         Recall = rep(NA, length(cms)),
                         F1 = rep(NA, length(cms)))

for (name in names(cms)) {
  label <- cms[[name]]
  
  comparison[comparison$model == name, ] <- comparison %>%
    filter(model == name) %>%
    mutate(Sensitivity = label$byClass[["Sensitivity"]],
           Specificity = label$byClass[["Specificity"]],
           Precision = label$byClass[["Precision"]],
           Recall = label$byClass[["Recall"]],
           F1 = label$byClass[["F1"]])
}

comparison %>%
  gather(x, y, Sensitivity:F1) %>%
  ggplot(aes(x = x, y = y, color = model)) +
    geom_jitter(width = 0.2, alpha = 0.5, size = 3)

}

set.seed(42)
index <- createDataPartition(bc_data$classes, p = 0.7, list = FALSE)
train_data <- bc_data[index, ]
test_data  <- bc_data[-index, ]

plot_cms(ctrl_sampling = sampling_list,
         target = "classes", #as.formula("classes~."),
         train = train_data,
          test = test_data)
```

