---
title: "Data Exploration and Auto reporting to doc, html, pdf, markdown format"
author: "Karim Mezhoud"
date: '`r Sys.Date()`'
output:
  html_document:
    number_sections: true
    fig_caption: true
    toc: true
    fig_width: 7
    fig_height: 6
    theme: cosmo
    highlight: tango
    code_folding: show #hide
    self_contained: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, error=FALSE, warning = FALSE)
knitr::opts_chunk$set(fig.width=12, fig.height=8) 
```



**The goal of this note is to report automatically  some clinical data of selected study from a public server**

# Load packages
```{r, message=FALSE}
library(cgdsr)
library(flextable)
library(dplyr)
```

## Connect to the public sever and load studies

```{r}
cgds <- cgdsr::CGDS("http://www.cbioportal.org/")
Studies<- cgdsr::getCancerStudies(cgds)
# look for Prostate studies
Prostate_studies <-
  Studies %>%
  filter(stringr::str_detect(name, "Prostate"))

#DT::datatable(Prostate_studies) %>%
 # DT::formatStyle( colnames(Prostate_studies), color = 'black')
knitr::kable(head(Prostate_studies), format = "html")
```

# Load clinical data of prostate studies
```{r}
clinicalData <- cgdsr::getClinicalData(cgds, "prad_tcga_all")
glimpse(clinicalData[1:20])
```

# Performe some exploration

## Median Age of `Reccured` and `DiseaseFree` for `Living` and `Deceased` patient

```{r}
clinicalData %>%
  group_by(DFS_STATUS, OS_STATUS, RACE) %>%
  summarise(Age = median(AGE))
```

## Which is the killer stage
```{r}
data <-
clinicalData %>%
  group_by(PATH_T_STAGE, OS_STATUS, DFS_STATUS) %>%
  summarise(Months = median(DFS_MONTHS))
data
```


```{r}
killer_stage <- 
clinicalData %>%
 group_by(CLIN_T_STAGE, PATH_T_STAGE, OS_STATUS, DFS_STATUS) %>%
  filter(OS_STATUS == 'DECEASED') %>%
  ungroup() %>%
  mutate(PATH_T_STAGE = as.factor(PATH_T_STAGE)) %>%
  count(PATH_T_STAGE) %>%
  top_n(1, PATH_T_STAGE)
 
killer_stage
```

# Built report
## Function to Built report

```{r}

flex_tbl <- 
  function( body ,  heads, colwidths, title){
    
    # define the body 
    ft <- flextable(body)
    # loop to add headers
    for( i in 1:length(heads)){
      ft <- add_header_row(ft, 
                           values = heads[[i]], 
                           colwidths = colwidth[[i]])
    }
    # Add Title
    ft <- add_header_lines(ft, title )
    
    # Set specific layout
    ft <- flextable::align(ft, i = 1, j = 1, part = c("header"), align = "center") 
    ft <- hline(ft, i = c(1, length(heads) + 1, length(heads) + 2),
                border = officer::fp_border(), part = "header")
    ft <- flextable::bold(ft, i = c(1, length(heads) + 2), bold = TRUE, part = "header")
     ft <- flextable::bold(ft, i = c(2, 3), j = c(1,4) , bold = TRUE, part = "header")
     ft <- flextable::color(ft, i = 3, j = c(3,4), color = 'red', part = "header")
    ft <- flextable::align(ft, part = "body", align = "left")
    ft <- fontsize(ft, size = 16, part = "all")
    ft <- padding(ft, padding = 10, part = "all")
    
    return(ft)
  }

```

## Formating data
```{r}
data <- clinicalData %>%
  group_by( PATH_T_STAGE, OS_STATUS, DFS_STATUS) %>%
  summarise(Months = median(DFS_MONTHS))

heads <- list(c("The killet stage", killer_stage[1], paste0("n = ", killer_stage[2])),
              c("Study", Prostate_studies[1,2])
                )

colwidth <- list(c(2,1,1), c(1,3))

flex_tbl(data, heads, colwidth, title = " Summary of the killer stage of Prostate Cancer")
```


