---
title: "Huxtable"
author: "Karim Mezhoud"
date: '`r Sys.Date()`'
output:
  html_document:
    number_sections: true
    fig_caption: true
    toc: true
    fig_width: 7
    fig_height: 6
    theme: cosmo
    highlight: tango
    code_folding: show #hide
    self_contained: no
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, error=FALSE)
knitr::opts_chunk$set(fig.width=12, fig.height=8) 
```


## getting start with a simple huxtable
```{r}
library(huxtable)
ht <- hux(
        Employee     = c('John Smith', 'Jane Doe', 'David Hugh-Jones'), 
        Salary       = c(50000L, 50000L, 40000L),
        add_colnames = TRUE
      )
## print as htlm
#print_html(ht)
## print as Latex
#print_latex(ht)

## print as markdown
ht
```

That is look like a datafarme but colnames is included in the first row of the dataframe.


## Changing the style of the hustable
```{r}
bold(ht)[1, ]          <- TRUE
bottom_border(ht)[1, ] <- 1
ht
```


## Tidyverse Syntaxe

```{r}
library(dplyr)
hux(
        Employee     = c('John Smith', 'Jane Doe', 'David Hugh-Jones'), 
        Salary       = c(50000, 50000, 40000),
        add_colnames = TRUE
      )                               %>%
      set_bold(1, 1:2, TRUE)          %>% 
      set_bottom_border(1, 1:2, 1)    %>%
      set_align(1:4, 2, 'right')      %>%
      set_number_format(2)            %>% 
      set_caption('Employee table')

```

## Changing content
```{r}
ht[3, 1] <- 'Jane Jones'
ht
```

## Adding content
```{r}
ht %>%
  cbind(c("Role", "Admin", "CEO", "Dogsbody")) %>%
  rbind( c("Yihui Xie", 100000, "CEO"))
```

## Editing content with `dplyr` way

```{r}
as_hux(mtcars) %>%
  add_colnames() %>%
  head

```

```{r}
as_hux(mtcars) %>%
      huxtable::add_rownames("Car")                       %>%
      slice(1:5)                                          %>% 
      select(Car, mpg, cyl, hp)                           %>% 
      arrange(hp)                                         %>% 
      filter(cyl > 4)                                     %>% 
      rename(MPG = mpg, Cylinders = cyl, Horsepower = hp) %>% 
      mutate(kml = MPG/2.82) %>%
      huxtable::add_colnames()  %>%
      theme_plain()
```


## Change widths
```{r}
width(ht) <- 0.4
col_width(ht) <- c(.7, .3)
ht
```

## Wrap  / stretched long content over multiple lines:
```{r}
ht_wrapped <- ht
ht_wrapped[5, 1] <- "David Arthur Shrimpton Hugh-Jones ejnskdkdjb  jnsdkdk,d jndjnd"
wrap(ht_wrapped) <- TRUE
ht_wrapped
```


## Merging cells
```{r}
#cars_ht <- as_huxtable(mtcars) %>% arrange(cyl)
  mtcars %>%
  as_huxtable() %>%
  huxtable::add_rownames("Car") %>%
  arrange(cyl) %>%
  select(1:4) %>%
 # mutate(cylinders = cars_ht$cyl) %>%
  mutate(cylinders = case_when(
    cyl == 4 ~ paste0("Four Cylinders"),
    cyl == 6 ~ paste0("Six Cylinders"),
    cyl == 8 ~ paste0("Eight Cylinders")
  )) %>%
  select(cylinders, everything()) %>%
  huxtable::add_colnames() %>%
  merge_cells(2:12, 1) %>% 
  merge_cells(13:19, 1) %>% 
  merge_cells(20:33, 1)


```

## Initiate empty hustable

### Head table
The Head tbale content the Title, and others informations about the report.
We set the head table with 4 columns and 4 rows.

```{r}
# header <- huxtable()
# head_description <- c("Title", "Description1", "Description2")
# head_value <- c(NA,NA, "Value2")
# header[seq_along(head_description), 1] <- head_description
# header[seq_along(head_value), 2] <- head_value
# 
# col_width(header[1,2]) <- 150
# bottom_border(header)[1, ] <- 1
# align(header)[1, 1] <- "center"
# bold(header)[1, ]    <- TRUE
# head <- merge_cells(header,1, 1:2)
# 
# theme_basic(header)
```



