---
title: "Pool MySQL Query"
author: "Karim Mezhoud"
date: '`r Sys.Date()`'
output:
  html_document:
    number_sections: true
    fig_caption: true
    toc: true
    fig_width: 7
    fig_height: 6
    theme: cosmo
    highlight: tango
    code_folding: show #hide
    self_contained: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, error=FALSE)
knitr::opts_chunk$set(fig.width=12, fig.height=8) 
```

This notebook aims to autogenerate specific format of report. We will start by using example of data from sql data base, do some processing and computing and that return a report.


# A simple example of reading dataset from sql database




# Install and manange local MySQL Database.

I found this [blog](https://programminghistorian.org/en/lessons/getting-started-with-mysql-using-r#installation-tips-for-a-mac) well exaplined to install and manage MySQL database.

[pool](https://db.rstudio.com/pool/) package makes easier to query SQL database.

# Example of amazon MySQL connection
```{r}
library(DBI)
library(dplyr)
library(RMySQL)
library(pool)

my_db <- dbPool(
  RMySQL::MySQL(), 
  dbname = "shinydemo",
  host = "shiny-demo.csa7qlmguqrf.us-east-1.rds.amazonaws.com",
  username = "guest",
  password = "guest"
)

dbListTables(my_db)
```


## Glimpse the Tables
```{r}

my_db %>% dplyr::tbl("City") %>% head(5)
```

```{r}
my_db %>% dplyr::tbl("Country") %>% head(5)
```


```{r}
my_db %>% dplyr::tbl("CountryLanguage") %>% head(5)
```

