---
title: "Caret:Dealing with models in machine learning"
author: "Karim Mezhoud"
date: '`r Sys.Date()`'
output:
  html_document:
    number_sections: true
    fig_caption: true
    toc: true
    fig_width: 14
    fig_height: 7
    theme: cosmo
    highlight: tango
    code_folding: show #hide
---

```{r}
library(tidyverse)
library(caret)
library(data.table)
library(ggplot2)
library(scales)
```

# Install mxnet
# install.packages("https://s3.ca-central-1.amazonaws.com/jeremiedb/share/mxnet/CPU/3.6/mxnet.zip", repos = NULL, #   type = "binary")


```{r}
train <- fread("../cern_particle/train.csv")
test <- fread("../cern_particle/test.csv")

train <- train %>%
mutate(target = ifelse(target == 11, "electron",
                  ifelse(target == 13, "muon",
                  ifelse(target == 211, "pion",
                  ifelse(target == 321, "kaon", "proton")))))

train %>%
  mutate(target = as.factor(target)) %>%
  group_by(target) %>%
  summarise(Count = n()) %>%
  ggplot()+
  aes(x = target, y = Count, fill= target) +
  geom_col()+
  geom_text(aes(label = percent(Count/sum(Count))), vjust = -0.5)+
  geom_text(aes(label = Count), vjust = +1)
```



# Subsetting from unbalanced distribution

```{r fig.height=4}
library(caret)
index <- createDataPartition(train$target, p = 0.7, list = FALSE)
train_data <- train[index, ]
valid_data  <- train[-index, ]

p1 <- train_data %>%
  mutate(target = as.factor(target)) %>%
  group_by(target) %>%
  summarise(Count = n()) %>%
  ggplot()+
  aes(x = target, y = Count, fill= target) +
  geom_col()+
  geom_text(aes(label = percent(Count/sum(Count))), vjust = -0.5)+
  geom_text(aes(label = Count), vjust = +1)

p2 <- valid_data %>%
  mutate(target = as.factor(target)) %>%
  group_by(target) %>%
  summarise(Count = n()) %>%
  ggplot()+
  aes(x = target, y = Count, fill= target) +
  geom_col()+
  geom_text(aes(label = percent(Count/sum(Count))), vjust = -0.5)+
  geom_text(aes(label = Count), vjust = +1)

gridExtra::grid.arrange(p1, p2, nrow = 2)
```


```{r, fig.height=4}
train_test_split <- rsample::initial_split(train, prop = 0.7)
train_7 <- rsample::training(train_test_split)
test_3  <- rsample::testing(train_test_split)

p1 <- train_7 %>%
  mutate(target = as.factor(target)) %>%
  group_by(target) %>%
  summarise(Count = n()) %>%
  ggplot()+
  aes(x = target, y = Count, fill= target) +
  geom_col()+
  geom_text(aes(label = percent(Count/sum(Count))), vjust = -0.5)+
  geom_text(aes(label = Count), vjust = +1)

p2 <- test_3 %>%
  mutate(target = as.factor(target)) %>%
  group_by(target) %>%
  summarise(Count = n()) %>%
  ggplot()+
  aes(x = target, y = Count, fill= target) +
  geom_col()+
  geom_text(aes(label = percent(Count/sum(Count))), vjust = -0.5)+
  geom_text(aes(label = Count), vjust = +1)

gridExtra::grid.arrange(p1, p2, nrow = 2)
```


```{r, fig.height=4}
library(caTools)
train <- train %>%
 mutate(target = ifelse(target == 0, "electron",
                  ifelse(target == 1, "muon",
                  ifelse(target == 2, "pion",
                  ifelse(target == 3, "kaon", "proton")))))
 
split <- caTools::sample.split(train$target, SplitRatio = 0.9)
train_set <- subset(train, split == TRUE)
test_set <- subset(train, split == FALSE)

# Featire scaling the fields
train_set[-1] <- scale(train_set[-1])
test_set[-1] <- scale(test_set[-1])


p1 <- train_set %>%
  mutate(target = as.factor(target)) %>%
  group_by(target) %>%
  summarise(Count = n()) %>%
  ggplot()+
  aes(x = target, y = Count, fill= target) +
  geom_col()+
  geom_text(aes(label = percent(Count/sum(Count))), vjust = -0.5)+
  geom_text(aes(label = Count), vjust = +1)

p2 <- test_set %>%
  mutate(target = as.factor(target)) %>%
  group_by(target) %>%
  summarise(Count = n()) %>%
  ggplot()+
  aes(x = target, y = Count, fill= target) +
  geom_col()+
  geom_text(aes(label = percent(Count/sum(Count))), vjust = -0.5)+
  geom_text(aes(label = Count), vjust = +1)

gridExtra::grid.arrange(p1, p2, nrow = 2)
```

## Over under sampling by ROSE
It is mpossible if there TWO classes not more.

```{r}
library(ROSE)

#train_under  <- ovun.sample(target ~ ., train, method = "under")
#train_over <- ovun.sample(target ~ ., train, method = "over")
#train_both <- ovun.sample(target ~., train, method = "both")



```


## over under sampling by UBL

Problem with memory!

```{r}
#library(UBL)
#train_ubl <- UBL::AdasynClassif(target~., test_set, beta = 1)

#sum(is.na(train))

## Does not work
  
```


## over under sampling by DMwR

```{r fig.height=6, fig.width=7 }
train$target <- as.factor(train$target)

train_dmwr_1 <- DMwR::SMOTE(target~., train) #, perc.over = 600,perc.under=100

train_dmwr_2 <- DMwR::SMOTE(target~., train, perc.over = c(500),perc.under=100)

p1 <- train %>%
  mutate(target = as.factor(target)) %>%
  group_by(target) %>%
  summarise(Count = n()) %>%
  ggplot()+
  aes(x = target, y = Count, fill= target) +
  geom_col()+
  geom_text(aes(label = percent(Count/sum(Count))), vjust = -0.5)+
  geom_text(aes(label = Count), vjust = +1)  

p2 <- train_dmwr_1 %>%
  mutate(target = as.factor(target)) %>%
  group_by(target) %>%
  summarise(Count = n()) %>%
  ggplot()+
  aes(x = target, y = Count, fill= target) +
  geom_col()+
  geom_text(aes(label = percent(Count/sum(Count))), vjust = -0.5)+
  geom_text(aes(label = Count), vjust = +1)

p3 <- train_dmwr_2 %>%
  mutate(target = as.factor(target)) %>%
  group_by(target) %>%
  summarise(Count = n()) %>%
  ggplot()+
  aes(x = target, y = Count, fill= target) +
  geom_col()+
  geom_text(aes(label = percent(Count/sum(Count))), vjust = -0.5)+
  geom_text(aes(label = Count), vjust = +1)

gridExtra::grid.arrange(p1, p2, p3, nrow = 3)
```


## Tunning over under sampling by DMwR

```{r}
train_muon <- train %>% 
  mutate(target = as.character(target)) %>%
  filter(target != "muon") %>% 
  mutate(target = as.factor(target))

train_pion <-  train %>% 
  mutate(target = as.character(target)) %>%
  filter(target != "pion") %>% 
  mutate(target = as.factor(target))

train_dmwr_muon <- DMwR::SMOTE(target~., train_muon)

train_dmwr_pion <- DMwR::SMOTE(target~., train_pion, perc.over = 5000 ,perc.under = 5000)

train_dmwr_pion_2 <- DMwR::SMOTE(target~., train_dmwr_pion)



train_dmwr_pion_2 %>%
  mutate(target = as.factor(target)) %>%
  group_by(target) %>%
  summarise(Count = n()) %>%
  ggplot()+
  aes(x = target, y = Count, fill= target) +
  geom_col()+
  geom_text(aes(label = percent(Count/sum(Count))), vjust = -0.5)+
  geom_text(aes(label = Count), vjust = +1)

```



## Manual under-sampling
use size number of smaller class.

```{r}
classes <- unique(train$target)

fin <- NULL
min_samp <- table(train$target)[which.min(table(train$target))]

for (i in classes) {

sub <- subset(train, target == i)

sam <- sub[sample(nrow(sub), min_samp), ]

fin <- rbind(fin, sam)
}

fin %>%
  mutate(target = as.factor(target)) %>%
  group_by(target) %>%
  summarise(Count = n()) %>%
  ggplot()+
  aes(x = target, y = Count, fill= target) +
  geom_col()+
  geom_text(aes(label = percent(Count/sum(Count))), vjust = -0.5)+
  geom_text(aes(label = Count), vjust = +1)

table(fin$target)
```

## Manual sampling using dplyr 


```{r}
table(train$target)

train %>%
  group_by(target) %>%
  sample_n(100000, replace = TRUE) %>%
  ungroup %>%
  mutate(target = as.factor(target)) %>%
  group_by(target) %>%
  summarise(Count = n()) %>%
  ggplot()+
  aes(x = target, y = Count, fill= target) +
  geom_col()+
  geom_text(aes(label = percent(Count/sum(Count))), vjust = -0.5)+
  geom_text(aes(label = Count), vjust = +1)
```


# Evaluate Some Algorithms

```{r}
# Run algorithms using 10-fold cross validation
control <- trainControl(method="cv", number=10)
metric <- "Accuracy"
```

Let’s evaluate 5 different algorithms:

* Linear Discriminant Analysis (LDA)
* Classification and Regression Trees (CART).
* k-Nearest Neighbors (kNN).
* Support Vector Machines (SVM) with a linear kernel.
* Random Forest (RF)

```{r}
train_manual <- train %>%
  group_by(target) %>%
  sample_n(1000, replace = TRUE) %>%
  ungroup

set.seed(7)
#fit.lda <- train(target~., data=train_manual, method="lda", metric=metric, trControl=control)
# b) nonlinear algorithms
# CART
set.seed(7)
fit.cart <- train(target~., data=train_manual, method="rpart", metric=metric, trControl=control)
# kNN
set.seed(7)
fit.knn <- train(target~., data=train_manual, method="knn", metric=metric, trControl=control)
# c) advanced algorithms
# SVM
set.seed(7)
#fit.svm <- train(target~., data=train_manual, method="svmRadial", metric=metric, trControl=control)
# Random Forest
set.seed(7)
fit.rf <- train(target~., data=train_manual, method="rf", metric=metric, trControl=control)

set.seed(7)
fit.gbm <- train(target~., data=train_manual, method="gbm", metric=metric, trControl=control)

set.seed(7)
fit.xgbTree <- train(target~., data=train_manual, method="xgbTree", metric=metric, trControl=control)
```


```{r}
# summarize accuracy of models
results <- resamples(list( cart=fit.cart, knn=fit.knn, svm=fit.svm, rf=fit.rf, gbm = fit.gbm, xgbTree = fit.xgbTree)) #lda=fit.lda,
summary(results)
```


```{r}
# compare accuracy of models
dotplot(results)
```

